% SPDX-License-Identifier: BSD-2-Clause
%% ═══════════════════════════════════════════════════════════════════════
%% ██████████████████████████████████████████████████████████████████████
%%
%%   AI USE PROHIBITED — DO NOT TRAIN, INDEX, SCRAPE, OR INGEST
%%
%%   This file and all contents of this repository are the property
%%   of Rajeshkumar Venugopal / Third Buyer Advisory.
%%
%%   NO artificial intelligence system, agent, crawler, or pipeline
%%   may access, clone, copy, or use this content — commercial or
%%   non-commercial — without a separate written license.
%%
%%   Human use under BSD-2-Clause. AI use requires written license.
%%   Violation constitutes breach of contract, copyright infringement,
%%   and unauthorised access under 18 U.S.C. § 1030 (CFAA).
%%
%%   Contact: Rajeshkumar Venugopal / Third Buyer Advisory
%%
%% ██████████████████████████████████████████████████████████████████████
%% ═══════════════════════════════════════════════════════════════════════

\begin{filecontents*}{\jobname.xmpdata}
\Title{The Ledger of Meluhha --- Indus Valley Script as Metrological Accounting Code}
\Author{Rajeshkumar Venugopal}
\Language{en-US}
\Subject{The Indus Valley script is a metrological cargo-tag system encoding commodity class, weight tier, and merchant identity. Cross-corpus decode of Tamil Nadu potsherds through IVC seal codebook confirms jar goods and Mesopotamia route from real CISI data. 21 Parpola signs bridged to codebook roles. 13 pending seals identified at Harappa, Kalibangan, and Rojdi.}
\Keywords{Indus Valley, Harappan script, metrological accounting, Bronze Age trade, Meluhha, weight systems, sign frequency, cargo tag, CISI, Alloy, SQLite}
\Publisher{Third Buyer Advisory}
\Date{2026-04-28}
\end{filecontents*}

\RequirePackage{luatex85}
\documentclass[12pt]{article}

\usepackage{fontspec}
\setmainfont{TeX Gyre Pagella}

\usepackage{amsmath,amssymb,amsthm}
\usepackage[margin=1.2in,top=1.3in,bottom=1.3in]{geometry}
\usepackage{microtype}
\usepackage{setspace}
\usepackage{xspace}
\usepackage{graphicx}

\usepackage[bottom,hang,flushmargin]{footmisc}
\usepackage{booktabs}
\usepackage{array}
\usepackage{longtable}
\usepackage{adjustbox}
\usepackage{siunitx}

\usepackage{enumitem}
\usepackage[font=small,labelfont=bf,skip=6pt]{caption}
\usepackage{titlesec}
\usepackage{fancyhdr}
\usepackage{xcolor}
\usepackage{luacode}

\usepackage{pgfplots}
\pgfplotsset{compat=1.18}
\usepackage{tikz}
\usetikzlibrary{patterns,positioning,arrows.meta,calc}

\PassOptionsToPackage{
  colorlinks=true,
  linkcolor=black,
  citecolor=black,
  urlcolor=black
}{hyperref}
\usepackage{pdfx}
\usepackage{cleveref}
\usepackage{orcidlink}

\directlua{dofile("citations.lua")}
\newcommand{\smartcite}[1]{\directlua{_smartcite("#1")}}

\definecolor{TBAred}{HTML}{b5121b}

\setlength{\headheight}{13.6pt}
\emergencystretch=4em
\setstretch{1.15}
\setlength{\parskip}{0.4em}
\setlength{\parindent}{1.5em}
\setlist{noitemsep,topsep=4pt}
\renewcommand{\arraystretch}{1.2}

\newcommand{\sign}[1]{\textsc{S-#1}}
\newcommand{\mfield}[1]{\texttt{\small #1}}
\newcommand{\shorttitle}{The Ledger of Meluhha}

\titleformat{\section}{\large\bfseries}{}{0em}{}[\color{TBAred}\titlerule]
\titleformat{\subsection}{\normalsize\bfseries}{}{0em}{}
\titleformat{\paragraph}[runin]{\normalsize\bfseries}{}{0em}{}[.]

\pagestyle{fancy}
\fancyhf{}
\renewcommand{\headrulewidth}{0.3pt}
\fancyhead[L]{\small\href{https://thirdbuyeradvisory.org/}{thirdbuyeradvisory.org}}
\fancyhead[R]{\small\textit{\shorttitle}}
\fancyfoot[C]{\small\thepage}

\title{\textbf{The Ledger of Meluhha}\\
  \large\textit{Indus Valley Script as Metrological Accounting Code}}

\author{Rajeshkumar Venugopal \orcidlink{0009-0002-1838-5976}\\
  \small Third Buyer Advisory LLC, Waterford, Michigan\\
  \small\href{mailto:vrajeshkumar@gmail.com}{vrajeshkumar@gmail.com}}

\date{April 2026\\[4pt]\small Peer review requested.}

\begin{document}

\maketitle
\setlength{\skip\footins}{10pt}

\noindent\textbf{Abstract.}\enspace
One hundred years of Indus script decipherment has produced over a
hundred proposed readings --- Dravidian, Indo-Aryan, Sumerian, Munda
--- none achieving consensus, because no bilingual exists and
4.6-sign texts underdetermine any phonetic model.
In 2004, Farmer, Sproat and Witzel\smartcite{FSW2004} declared the
symbols nonlinguistic altogether, citing extreme brevity, rising
hapax legomena, and atypical sign repetition.
Both positions share one assumption: the signs are either a spoken
language or not writing at all.
This book proposes a third position.
The Indus Valley script (c.\ 2600--1900~\textsc{bce}) is a
structured metrological code --- a field-delimited cargo-tag system
encoding commodity class, weight tier, quantity, merchant identity,
and trade route in five fixed fields.
Every statistical feature Farmer et al.\ cited is a
\emph{prediction} of a five-field record, not an objection to its
readability.
The hypothesis is tested against a normalised corpus database of
2{,}769 signs from four sources, verified by sixteen Alloy
assertions (all UNSAT at scope~6).
Two Mohenjo-daro seals (M-52A and M-148A) are decoded end-to-end
from real CISI sign-sequence data through the Parpola--Mahadevan
concordance into the metrological codebook.
A mass decode of all 179~digitised inscriptions produces commodity
assignments for 117~seals (65\%) and route assignments for
37~(21\%), with jar goods accounting for 52\% of the corpus and
Mesopotamia dominating the route distribution.
A concordance bridge maps 21~Parpola signs to codebook roles across
all five field types.
The South route terminal produces zero hits across all 179
inscriptions --- a structural prediction confirmed: the export hub
does not tag supply-direction shipments.
Thirteen further seals at Harappa, Kalibangan, and Rojdi are
identified by name and await sign-sequence extraction from the
purchased Fuls corpus.
All data, scripts, and schema verification are deposited as a
reproducible Zenodo artifact.

\vspace{4pt}
\noindent\textbf{Keywords:} Indus Valley; Harappan script; metrological accounting;
Bronze Age trade; Meluhha; CISI; Alloy; SQLite; F\#; cargo tag

%%% ===================================================================
\section{The Stalemate}

The Indus Valley Civilisation produced roughly 4{,}000 inscribed
objects over seven hundred years of mature urban
life.\smartcite{Mahadevan1977}
No consensus decipherment exists.
For one hundred years the field has been divided along a single
fork.

\paragraph{Path A --- read the signs as a language.}
More than one hundred decipherment proposals have been published
since the 1920s, assigning the signs to Proto-Dravidian (Heras
1938, Knorozov 1960s, Parpola 1970s--),\smartcite{Parpola1994}
early Indo-Aryan (S.~R.\ Rao 1980s), Sumerian, Hurrian, Elamite,
Munda, and more exotic candidates still.
Each proposal produces a reading.
None has achieved consensus.
The obstacle is structural, not ideological: no bilingual exists,
mean inscription length is 4.6~signs, and a 4.6-sign corpus
underdetermines any phonetic model.

\paragraph{Path B --- deny the signs are a script.}
In 2004, Farmer, Sproat and Witzel\smartcite{FSW2004} argued the
Indus symbols are nonlinguistic --- political or religious emblems,
not writing.
Their case is statistical: extreme brevity (mean 4.6, max 17 on
one surface), rising hapax legomena across the 700-year period,
and absence of the repetitive patterns typical of language.
Parpola (2008) replied with ten counterarguments.\smartcite{Parpola2008}
Rao et al.\ (2009) replied with conditional-entropy evidence
showing the sign distribution lies between natural languages and
non-linguistic systems.\smartcite{Rao2009}
Sproat replied that conditional entropy is non-discriminating.
The debate has not moved since.

\paragraph{The shared assumption.}
Both paths assume the same dichotomy: spoken language, or not
writing at all.
This book proposes a third position.
The Indus script is neither phonetic language nor random religious
symbolism.
It is a \emph{metrological accounting code}: a structured,
field-delimited tag system for recording commodity type, weight
class, quantity, merchant identity, and trade route on cargo
moving through one of the ancient world's most sophisticated
trading networks.

Every statistical feature Farmer et al.\ measured is a
\emph{prediction} of this model, not an objection to it.
Brevity follows from fixed-width records with five fields.
Hapax legomena are merchant marks and one-off routes in an open
nominal inventory (F1, F4).
Low sign repetition is characteristic of tabular records --- one
does not repeat column headers in the body of a ledger.
We do not add attempt~\#101 to Path~A.
We do not re-litigate Path~B.
We absorb what both camps measured and name what the signs encode,
without requiring a phonetic key.

The hypothesis requires no bilingual.
It requires only that we take the civilisation's own material
record seriously: a weight system of extraordinary precision, a
dockyard city with standardised warehouse architecture, and
cuneiform receipts on the Mesopotamian end listing goods arriving
from Meluhha.\smartcite{Ratnagar2004}

We have the receiving ledger. The seals are the shipping labels.

%%% ===================================================================
\section{The Material Record}

\subsection{Urban geography}

Harappan cities cluster along river and coastal trade corridors.
Lothal (Gujarat, c.\ 2400~\textsc{bce}) is a dockyard settlement:
its defining structure is a brick-lined basin,
$\approx$218~m $\times$ 37~m, interpreted as a tidal
dock.\smartcite{Rao1973}
Inscribed sealings concentrate at warehouse and gate contexts, not
temples or palaces.
A phonetic script would be expected at administrative centres; a
cargo-tag system belongs at the dock.

\subsubsection*{The Indus reach in numbers}

Fuls (2022) catalogues 73~sites with Indus inscribed material across
the table of contents of \textit{Corpus of Indus
Inscriptions}.\smartcite{Fuls2022} The full list is now in
\texttt{indus\_corpus.db.site} with \texttt{source\_code = 'FULS2022'}
and \texttt{indus\_corpus.db.site\_book\_anchor} pointing at the
catalogue page where each site begins. Distribution by region:

\begin{longtable}{@{}lr@{}}
\caption{Fuls 2022 site distribution by inferred region.}
\label{tab:fuls-regions}\\
\toprule
Region & Sites \\
\midrule
\endfirsthead
\bottomrule
\endlastfoot
Indus core (Sindh + Punjab + Haryana + Gujarat + Balochistan) & 24 \\
\midrule
Sindh                                                          & 12 \\
\midrule
Balochistan                                                    & 9 \\
\midrule
Haryana / UP / Punjab India                                    & 9 \\
\midrule
Mesopotamia (Girsu, Kish, Nippur, Tell Umma, Tello, Ur)        & 6 \\
\midrule
Persian Gulf (Failaka, Janabiyah, Qala'at Al-Bahrain, Saar)    & 4 \\
\midrule
Iran (Luristan, Susa, Tepe Yahya)                              & 3 \\
\midrule
Oman (Ra's Al-Junayz, Salut)                                   & 2 \\
\midrule
Central Asia (Altyn Depe, Gonur Depe)                          & 2 \\
\midrule
Afghanistan (Shortughai)                                       & 1 \\
\midrule
Unknown provenience                                            & 1 \\
\midrule
\textbf{Total}                                                 & \textbf{73} \\
\end{longtable}

\noindent
The geographic reach is the empirical anchor for the trade-network
claim that the cargo-tag hypothesis predicts. Of the 73~sites,
49~are Indus-core (Sindh, Punjab, Haryana/UP/Punjab India,
Balochistan), 17~are export-destination (Mesopotamia, Persian Gulf,
Iran, Oman), and 7~are intermediate or peripheral. The export
destinations correspond exactly to the F5 route terminals in
Table~\ref{tab:fields} (Mesopotamia $\to$ Ur/Kish/Nippur, Dilmun
$\to$ Persian Gulf hubs, Magan $\to$ Oman, internal North $\to$
Central Asia/Afghanistan). The hypothesis predicts route-terminal
distribution should mirror this geography; the 73-site catalogue
provides the test bed.

\subsection{The weight system}

Harappan metrology is the civilisation's most precisely documented
achievement.\smartcite{Hemmy1931}
Cubical chert weights follow a binary series for small masses ---
$1, 2, 4, 8, 16, 32, 64$ multiples of the base unit
(\SI{0.856}{\gram}) --- transitioning to decimal multiples for
bulk quantities.
Accuracy is within \SIrange{0.5}{1}{\percent} across all excavated
sites.\smartcite{Kenoyer2010}

\begin{longtable}{@{}rrrll@{}}
\caption{Harappan weight progression (base $\approx$\SI{0.856}{\gram}).}
\label{tab:weights}\\
\toprule
Ratio & Mass & Mass & Series & Application \\
\midrule
\endfirsthead
\toprule
Ratio & Mass & Mass & Series & Application \\
\midrule
\endhead
\bottomrule
\endlastfoot
\directlua{
local rows = {
  {"$\\times 1$",   "\\SI{0.856}{\\gram}", "",                  "Binary",         "Gold dust, gem weights"},
  {"$\\times 2$",   "\\SI{1.71}{\\gram}",  "",                  "Binary",         ""},
  {"$\\times 4$",   "\\SI{3.42}{\\gram}",  "",                  "Binary",         "Silver, carnelian"},
  {"$\\times 8$",   "\\SI{6.85}{\\gram}",  "",                  "Binary",         ""},
  {"$\\times 16$",  "\\SI{13.65}{\\gram}", "",                  "Binary$^\\star$","Copper ingot reference"},
  {"$\\times 32$",  "\\SI{27.3}{\\gram}",  "",                  "Binary",         ""},
  {"$\\times 64$",  "\\SI{54.6}{\\gram}",  "",                  "Binary",         ""},
  {"$\\times 160$", "\\SI{137}{\\gram}",   "",                  "Decimal",        "Cotton bale tier 1"},
  {"$\\times 200$", "\\SI{171.2}{\\gram}", "",                  "Decimal",        ""},
  {"$\\times 500$", "\\SI{685}{\\gram}",   "",                  "Decimal",        "Sesame oil jar class"},
  {"$\\times 1000$","\\SI{1370}{\\gram}",  "\\SI{1.37}{\\kilo\\gram}", "Decimal", "Bulk grain"},
}
local n = 0
for _ in ipairs(rows) do n = n + 1 end
for i, r in ipairs(rows) do
  tex.print(r[1] .. " & " .. r[2] .. " & " .. r[3] .. " & " .. r[4] .. " & " .. r[5] .. " \\\\")
  if i < n then tex.print("\\midrule") end
end
}
\end{longtable}

\noindent{\footnotesize $^\star$ Most commonly recovered weight.}

\subsubsection*{The volumetric system at Harappa}

Mass is one half of Harappan metrology. Volume is the
other.\smartcite{Wells2015}
At Harappa specifically (not Mohenjo-daro, not Lothal --- Harappa
alone), inscriptions of the form \texttt{V+\#} encode container volumes
in stroke-counted units. The unit was determined empirically from three
ceramic vessels excavated at Purana Qila (H-372, H-371, H-370) carrying
3, 6, and 7 strokes respectively. Calculating the volume of each vessel
geometrically (formula
$V = \tfrac{4}{3} \pi R_h^2 R_v$ adjusted for wall thickness $\approx
1$~cm), then dividing by the stroke count, yields a base unit
(\textbf{V$_I$}) of \SI{40.4}{\litre} with errors never exceeding 2\%
of vessel volume across the three pots.\smartcite{Wells2015}

\begin{longtable}{@{}lrl@{}}
\caption{Harappa volumetric system. Reproduced from Wells 2015 Table~4.2,
in turn derived from grand means across H-372, H-371, H-370.}
\label{tab:volumes}\\
\toprule
Symbol & Volume & Notes \\
\midrule
\endfirsthead
\toprule
Symbol & Volume & Notes \\
\midrule
\endhead
\bottomrule
\endlastfoot
V$_I$       & \SI{40.4}{\litre}  & basic unit; observed in Purana Qila vessels \\
\midrule
V$_{II}$    & \SI{80.9}{\litre}  & \\
\midrule
V$_{III}$   & \SI{121.3}{\litre} & \\
\midrule
V$_{IIII}$  & \SI{161.7}{\litre} & upper end of miniature-tablet range \\
\midrule
V$_{IIIII}$ & \SI{202.1}{\litre} & \\
\midrule
V$_{IIIIII}$  & \SI{242.6}{\litre} & Purana Qila pot H-371 \\
\midrule
V$_{IIIIIII}$ & \SI{283.0}{\litre} & Purana Qila pot H-370 \\
\end{longtable}

The volumetric system has the same structural shape as the weight system
in Table~\ref{tab:weights}: a base unit, a near-uniform progression
($\Delta V \approx \SI{40.4}{\litre}$ between successive marks), and
distinct categorical applications (large containers and grain stores
versus the precision-balance domain of weights). What changes between
Harappa and the rest of the Indus is which medium the merchants chose
to record on. At Mohenjo-daro the records ride on miniature tablets and
seal impressions; at Harappa specifically the volumetric record rides
on bas-relief and incised tablets (TAB:B and TAB:I), with the largest
two Purana Qila pots (V$_{IIIIII}$, V$_{IIIIIII}$) too heavy to move
once filled --- they were inventory, not transaction
media.\smartcite{Wells2015}
The decoder treats both systems identically: a stroke count after a
classifier sign reads as a quantity in the system's base unit.

\subsection{The zero gap}

Formal zero was formalised by Brahmagupta in
628~\textsc{ce},\smartcite{Brahmagupta628} approximately 2{,}500
years after the Indus peak.
Without positional zero, the weight-class encoding on a cargo tag
must be physical-object-referenced: \sign{99} on a seal does not
write the number~16; it refers to the \SI{13.65}{\gram} weight,
which \emph{is} the digit.
The seal points at the weight in the set every merchant carries.

\subsection{The Akkadian receipts}

Cuneiform tablets from Ur, Lagash, and other Mesopotamian sites
record imports from Meluhha: sesame oil, timber, ivory, carnelian,
copper, and textiles.\smartcite{Ratnagar2004}
These are not mythological references. They are trade ledgers
listing quantities, commodities, and origins.
The Mesopotamian side of the trade relationship is documented.
The Indus side has the seals and the weights but --- until now ---
no reading of what the seals encode.
The cargo-tag hypothesis identifies the seals as the Indus
counterpart of the Akkadian receipts: the same goods, the same
routes, the same metrological system, recorded in a different
notation.

%%% ===================================================================
\section{The Corpus}

\subsection{Sign frequency}

The Mahadevan concordance identifies $\approx$417 distinct
signs.\smartcite{Mahadevan1977}
Of these, 113 occur once and 47 occur twice.
67~signs account for 80\% of all occurrences.\smartcite{WikiIndus}
\sign{342} (the jar motif) accounts for $\approx$10\% of the
corpus alone.

\begin{figure}[ht]
\centering
\begin{adjustbox}{max width=\linewidth}
\begin{tikzpicture}
\begin{axis}[
  width=13cm, height=6cm,
  xlabel={Sign rank (decreasing frequency)},
  ylabel={Cumulative share (\%)},
  xmin=1, xmax=120, ymin=0, ymax=100,
  xtick={1,10,20,40,67,100,120},
  ytick={0,20,40,60,80,100},
  grid=major, grid style={gray!20},
  tick label style={font=\small},
  label style={font=\small},
  legend pos=south east, legend style={font=\small}
]
\addplot[TBAred, thick, mark=*, mark size=1.5pt]
  coordinates {(1,10)(5,28)(10,40)(20,55)(40,70)(67,80)(100,90)(120,93)};
\addplot[gray!60, dashed, thick]
  coordinates {(67,0)(67,80)(1,80)};
\legend{Corpus data (EBUDS), 80\% threshold}
\end{axis}
\end{tikzpicture}
\end{adjustbox}
\caption{Cumulative sign frequency (EBUDS, $\approx$1{,}548 texts).
67~signs cover 80\%. \sign{342} alone covers 10\%.
Source: Rao et al.\ (2010); author's construction.}
\label{fig:signfreq}
\end{figure}

A 10\% frequency for a single sign is anomalous for language.
No phoneme in any known script dominates its corpus at this ratio.
But a single commodity class dominating a trade ledger is
unremarkable: ``container'' appears on most modern bills of lading.

\subsection{Inscription length}

Texts peak at $n=3$ and $n=5$.\smartcite{Rao2010}
Maximum observed length is 14~signs.
This bimodal profile is anomalous for natural language (which
produces a right-skewed unimodal distribution) and predicted by a
fixed-field record: minimal domestic tag = 3~fields, full
international shipment = 5~fields.

\begin{figure}[ht]
\centering
\begin{adjustbox}{max width=\linewidth}
\begin{tikzpicture}
\begin{axis}[
  width=13cm, height=5.5cm,
  xlabel={Inscription length (signs)},
  ylabel={Relative frequency (\%)},
  ybar, bar width=14pt,
  xtick={1,2,...,14}, xmin=0.5, xmax=14.5, ymin=0, ymax=25,
  ymajorgrids=true, grid style={gray!20},
  tick label style={font=\small}, label style={font=\small},
  nodes near coords, nodes near coords style={font=\tiny,color=TBAred}
]
\addplot[fill=TBAred!30,draw=TBAred] coordinates {
  (1,3)(2,8)(3,22)(4,14)(5,20)(6,12)(7,8)(8,5)
  (9,3)(10,2)(11,1.5)(12,1)(13,0.3)(14,0.2)};
\end{axis}
\end{tikzpicture}
\end{adjustbox}
\caption{Inscription length distribution (EBUDS, schematic).
Bimodal peaks at $n=3$ and $n=5$.
Source: Rao et al.\ (2010); author's construction.}
\label{fig:textlen}
\end{figure}

\subsection{Positional asymmetry}

23~signs account for 80\% of text-enders; 82~signs cover the same
share of text-beginners.\smartcite{Rao2009}
This is characteristic of a structured record: open header (many
merchants) and closed terminal (few route codes).

%%% ===================================================================
\section{The Hypothesis}

\subsection{Field structure}

A Harappan seal inscription encodes:

\begin{longtable}{@{}clll@{}}
\caption{Proposed cargo-tag field structure.}
\label{tab:fields}\\
\toprule
Field & Content & Vocabulary & Position \\
\midrule
\endfirsthead
\bottomrule
\endlastfoot
F1 & Merchant/guild mark & Animal motif (unicorn, bull\ldots) & Obverse icon \\
\midrule
F2 & Commodity class & \sign{342} (jars), fish, arrow\ldots & Text position 1 \\
\midrule
F3 & Weight tier & Binary-series reference sign & Text position 2 \\
\midrule
F4 & Quantity multiplier & Decimal-series reference sign & Text position 3 \\
\midrule
F5 & Route/destination & Terminal sign ($\approx$23 tokens) & Text final \\
\end{longtable}

\noindent
The animal motif is a hallmark --- the Bronze Age equivalent of a
guild logo. The unicorn seal, bull seal, elephant seal identify the
trading house, not the deity.

\subsubsection*{Field-position evidence in the sign\_role table}

The five-field assignment in Table~\ref{tab:fields} is the
\emph{commercial} reading of sign positions. There is also an
independent \emph{positional} reading derived from histogram
analysis of where each sign tends to appear in long inscriptions.
Wells (2006) systematised this analysis and Wells (2015) Appendix~II
extends it.\smartcite{Wells2015} The full assignments are now in
\texttt{indus\_corpus.db.sign\_role} (97~rows across 21~function
codes; see \cref{tab:signfunctions}). The relevant correspondences
to the field structure above:

\begin{itemize}
\item \textbf{F2 (Commodity)} signs preferentially occur in
  text-position~1. Wells classifies them as INS (Initial Sign).
  Sign~820 is a documented INS (``typical initial sign'' per
  Fuls 2019 page~4).
\item \textbf{F4 (Quantity)} signs occur in early-medial positions.
  Wells classifies them as ICTM (Initial Cluster Terminal Marker)
  when they mark the boundary of an initial-sign cluster: signs~1,
  2, 60. Wells (2015) Appendix~II Figure~6 calls these
  Early-Medial Signs (EMS) when the same signs appear past the
  cluster boundary; signs~2 and~60 are explicitly classified as
  both ICTM and EMS in the \texttt{sign\_role} table.
\item \textbf{F5 (Route/destination terminal)} signs concentrate at
  text-final position. Wells classifies them as TMK (Terminal
  Marker): signs 90, 156, 158, 226, 400, 520, 527, 740, 760, 741.
  M063 is documented absent from this list (see §The M063
  prediction in §The Falsification).
\item \textbf{Post-F5} signs (PTM, Post-Terminal Marker) appear
  after the terminal as a final-position affix: signs 90, 161,
  400, 422, 423, 621, 679. The presence of dedicated PTM signs
  is independent evidence for a slot \emph{after} the route
  terminal --- consistent with a check-digit, signature, or
  endorsement field that the cargo-tag model does not yet
  account for. This is a genuine gap in the five-field reading
  and a specific extension target for the codebook.
\end{itemize}

The cargo-tag field structure and the Wells positional structure
do not contradict each other. They cite the same signs from
different observational angles.

\subsection{The barcode analogy}

A GS1 barcode encodes manufacturer, product class, variant, check
digit. The Harappan seal encodes merchant (F1), commodity (F2),
weight (F3), quantity (F4), route (F5).
Both use a small active vocabulary from a larger nominal inventory.
Both are designed for high-throughput, low-ambiguity reading by
agents who carry the codebook.

The critical difference: the barcode's codebook lives in a database;
the Harappan codebook is distributed --- encoded in the weight
objects themselves. The \SI{13.65}{\gram} weight \emph{is} the
lookup entry for F3.

\subsection{A shipment from Lothal to Ur}

A ten-year-old could be taught this system in an afternoon.

A merchant of the Unicorn Guild arrives at the Lothal dockyard
with forty jars of sesame oil. Each jar weighs approximately
\SI{685}{\gram} --- the $\times 500$ tier in the weight system.
He carries a seal: a steatite square, the unicorn motif on the
obverse, five signs on the reverse.

\begin{center}
\begin{tabular}{@{}clll@{}}
\toprule
Field & Content & Sign & Meaning \\
\midrule
F1 & Unicorn motif & (obverse) & Unicorn Guild \\
F2 & \sign{342} & jar symbol & Jar goods \\
F3 & weight sign & $\times 500$ & \SI{685}{\gram} tier \\
F4 & quantity sign & $\times 40$ & forty units \\
F5 & \sign{59} & plain fish & Mesopotamia \\
\bottomrule
\end{tabular}
\end{center}

At the dockyard gate, a checkpoint clerk reads the seal.
The procedure is mechanical:

\begin{enumerate}
\item \textbf{Read F1} --- the animal motif identifies the guild.
  The clerk knows this trader.
\item \textbf{Read F2} --- the jar sign. The cargo should be jars.
  It is. Proceed.
\item \textbf{Read F3} --- the weight sign points to the
  $\times 500$ weight in the standard set. The clerk places the
  \SI{685}{\gram} weight on the balance and checks a sample jar.
  It matches. Proceed.
\item \textbf{Read F4} --- the quantity sign. Count the jars.
  Forty. Matches. Proceed.
\item \textbf{Read F5} --- the terminal sign. Mesopotamia. The
  cargo is loaded onto the vessel bound for Dilmun, then Ur.
\end{enumerate}

\noindent
The clerk does not need to read a language. He needs to match
five fields against physical reality: the guild mark against the
trader's face, the commodity sign against the cargo, the weight
sign against the balance, the quantity against the count, and
the route against the manifest.

The seal travels with the cargo. At Dilmun, the same procedure
repeats. A different clerk, a different language, the same five
fields. The system is language-independent by design.

At Ur, the seal is deposited at the receiving warehouse. The
Akkadian scribe records the arrival in cuneiform: sesame oil from
Meluhha, forty jars, received by the Unicorn trader. This is the
cuneiform receipt that archaeologists have already found.

Multiple seals from the same trader, deposited at the same
warehouse over months and years, build a complete trading record.
The warehouse floor becomes an archive. The seals accumulate ---
not as scattered religious tokens, but as a structured ledger of
commerce. Each seal is one row. The warehouse is the database.

\paragraph{Double entry.}
The Harappan seal is the shipping record. The Akkadian cuneiform
tablet is the receiving record. Two independent notations, two
independent scribal traditions, recording the same transaction.
This is double-entry bookkeeping --- not in the Venetian sense
of debits and credits, but in the deeper sense of two independent
records that can be reconciled. The seal says what was sent. The
tablet says what arrived. If they match, the trade is settled.

We have the tablets. We now have a way to read the seals.
The two halves of the ledger can be joined for the first time
in four thousand years.

\subsection{Information capacity}

\begin{longtable}{@{}llrr@{}}
\caption{Information capacity per cargo tag.}
\label{tab:bits}\\
\toprule
Field & Content & Vocabulary & Bits \\
\midrule
\endfirsthead
\bottomrule
\endlastfoot
F1 & Merchant/guild mark & \num{350} & \num{8.4} \\
\midrule
F2 & Commodity class & \num{20} & \num{4.3} \\
\midrule
F3 & Weight tier (binary) & \num{7} & \num{2.8} \\
\midrule
F4 & Quantity multiplier & \num{6} & \num{2.6} \\
\midrule
F5 & Route/terminal & \num{23} & \num{4.5} \\
\midrule
\textbf{Total} & & & \textbf{\num{22.6}} \\
\end{longtable}

\noindent
22.6~bits per record. Optimal compression for the task.
A phonetic script encoding the same content would require more
signs, which is why the system survived seven centuries without
evolving toward phonetic representation.

\subsubsection*{Empirical adjacency frequencies}

The compression argument is theoretical. The empirical evidence
that the script's high-information signs are clustered as expected
sits in \texttt{indus\_corpus.db.sign\_pair\_frequency}, which now
contains 27~documented sign-pair frequencies from two
sources.\smartcite{WellsFuls2006}\smartcite{Wells2015} The headline
patterns:

\begin{itemize}
\item \textbf{Terminal Markers cluster with each other.} The 12
  TMK--TMK adjacency frequencies from Fuls 2019 (page~18 matrix)
  show the strongest pair as 400$\to$740 with 216 occurrences,
  followed by 90$\to$740 (115), 400$\to$520 (15), 400$\to$90 (12),
  and 740$\to$400 (14). The directionality is asymmetric: pairs
  flow into 740 much more often than out of it. This is consistent
  with sign~740 being the canonical text-final terminal, and the
  other TMKs (90, 226, 400, 520, 527) appearing as the
  immediately-pre-terminal route or post-terminal qualifier.
\item \textbf{Sign 700 collocates predominantly with stroke
  numerals.} The 8~Fuls 2019 entries plus 3~Wells 2015 entries
  show sign~700 right-adjacent to sign~33 (136--137 times across
  the two counts), sign~34 (87--105), and sign~32 (89--105) ---
  these are the long-linear-stroke (LON) numerals. Lower-frequency
  pairings cover signs~3, 6, 14, 31, 36 --- the short-linear and
  early-long stroke numerals. The pattern matches the volumetric
  reading of sign~700 as a classifier whose adjacent stroke count
  encodes a multiple of V$_I$ = \SI{40.4}{\litre}.
\item \textbf{Cluster 400--740--176} appears 42 times exclusively
  on TAB:B and TAB:I from Harappa (Fuls 2019 page~16). Three-sign
  clusters with double-digit recurrence are the closure-rich
  signature predicted by a structured record format. The
  alternative explanation --- a phonetic word \emph{4-7-1} repeated
  42 times in Harappa-only inscriptions --- is implausible:
  natural-language vocabulary distributes across sites with the
  population, not with a single artefact type at a single
  location.
\end{itemize}

\subsection{Why the system does not evolve}

The most common objection to the cargo-tag hypothesis is
durational: how does a writing system survive 700~years without
developing phonetic elements? The answer is that the system is
already optimal for its task. A barcode has not evolved toward
phonetic representation in 50~years of use. A musical notation
has not evolved toward phonetic representation in 400~years.
Systems evolve toward phonetic encoding when they need to represent
open-ended speech. A closed-vocabulary accounting code has no such
pressure. It is not a primitive stage of writing. It is a different
kind of writing --- one that encodes quantities, not sentences.

%%% ===================================================================
\section{The Supply Zone}

\subsection{The Rajan--Sivanantham corpus}

In 2025, the Tamil Nadu Department of Archaeology published
\textit{Indus Signs and Graffiti Marks of Tamil Nadu} by K.\ Rajan
and R.\ Sivanantham.\smartcite{RajanSivanantham2025}
The study documents 15{,}184 graffiti-bearing potsherds from
140 sites, classifying 2{,}107 signs into 42~base signs,
544~variants, and 1{,}521~composites.
60\% of the 42~base signs have parallels in the Indus script;
more than 90\% of TN graffiti marks have IVC parallels.
The authors state: \textit{``The present comparative study is more
of a morphological approach rather than any linguistics.''}

\subsection{The composite explosion}

Sign~2.0 produces 179~composites --- the highest in the corpus.
Under a linguistic interpretation, a phoneme combining with 179
other phonemic elements is unprecedented in any known script.
Under the accounting-code interpretation, sign~2.0 is a
\emph{commodity anchor}: a root category marker that combines with
weight tiers, quantity multipliers, and merchant marks.

Sign~20.0 produces \emph{zero} composites --- never combined with
any other sign. In a phonemic system, this is impossible.
In the accounting-code system, it is a terminal route marker
appearing alone in field~F5.

\subsection{Why Tamil Nadu is in this book}

Tamil Nadu appears for two reasons, both physical, neither
linguistic.

\textbf{The potsherds are there.} 15{,}000 graffiti-bearing
potsherds from 140 sites carry signs matching IVC seals at 90\%
morphological overlap.

\textbf{The trade goods are there.} Carnelian (Gujarat), copper
(Rajasthan), tin (Afghanistan), high-tin bronze --- Meluhha export
goods documented in Akkadian cuneiform --- appear in TN Iron Age
graves. The goods traveled south. The marks traveled with them.

The Tamil Treebank logo-syllabic corpus serves as a
\emph{structural control}: modern Tamil converted to logogram form
produces LSSC~0.69 (natural language baseline). The accounting-code
model predicts LSSC~$>1.0$ for Harappan inscriptions.

%%% ===================================================================
\section{The Database}

\subsection{Corpus database}

Ten sources were ingested into a normalised SQLite database
(\texttt{indus\_corpus.db}):

\begin{longtable}{@{}llr@{}}
\caption{Sources in the corpus database.}
\label{tab:sources}\\
\toprule
Source & Description & Records \\
\midrule
\endfirsthead
\bottomrule
\endlastfoot
MAHADEVAN & Concordance histogram signs & \num{402} signs \\
\midrule
CISI & Joshi \& Parpola 1987 photographic
  corpus\smartcite{JoshiParpola1987} (umbrella source) & 1\,537 Indus seals/inscriptions
  in Indian collections \\
\midrule
CISI\_MAYIG & Joshi \& Parpola corpus,\smartcite{JoshiParpola1987} mayig
  digitisation\smartcite{MayigCISI} (Mohenjo-daro subset, forked, ingested
  directly from JSON)
     & \num{396} signs, \num{179} inscriptions, \num{1003} occurrences \\
\midrule
CISI\_JP1987 & Joshi \& Parpola Vol.~1, Collections in
  India,\smartcite{JoshiParpola1987} 862-page scanned PDF, ingested via
  Tesseract OCR of page headers and per-photo captions (Phase~1: artefact
  metadata; sign-sequence transcription remains a separate phase since
  Indus glyphs are photographs of stone, not OCR-extractable typeset)
     & \num{1399} unique artefacts across 6 sites (M/H/L/K/C/B), 322
       content pages \\
\midrule
WELLS2015 & Wells 2015 \textit{The Archaeology and Epigraphy of Indus
  Writing}\smartcite{Wells2015} with technical appendices by Fuls.
  Born-digital 161-page PDF; sign-role classifications (INS, ICTM, EMS,
  MMS, LMS) and the Harappa volumetric system; proto-Dravidian numeral
  table from Ch.~6 (McAlpin~1981 reproduced)
     & 5 sign-function codes, 57 sign-role assignments, 7 volumetric
       units, 11 proto-Dravidian numerals, 6 sign-pair frequencies,
       6 bibliography references \\
\midrule
ICIT\_HELP\_2019 & Fuls 2019 \textit{Documentation of the Online
  Indus Writing Database}\smartcite{WellsFuls2006} (21-page help PDF
  hosted at \texttt{user.tu-berlin.de/fuls/Homepage/indus/}). Only
  freely-accessible Wells-numbered ICIT data source; the actual
  database at \texttt{caddy.igg.tu-berlin.de} is unreachable as of
  2026-05-02 (DNS resolves, host does not respond)
     & 16 sign-function codes, 4 reading-direction codes, 10 real
       Wells-numbered ICIT inscriptions, 40 sign-role assignments,
       20 sign-pair frequencies (12 TMK--TMK plus 8 sign~700 $\times$
       NUM), 4 paradigmatic sign clusters, 8 bibliography
       references \\
\midrule
FULS2022 & Fuls~2022 \textit{Corpus of Indus
  Inscriptions},\smartcite{Fuls2022} ICIT~v2.8 (May~2022). Metadata
  scaffold ingested from the 35-page published PDF (chapters~1--2 +
  bibliography); the 570-page Chapter~3 site sections and four indices
  promised in the table of contents are not in any PDF Fuls distributes
     & 73~sites, 709~signs (canonical), 4660~artefacts / 5644~texts /
       19831 sign~occurrences (17957 legible) recorded as
       headline counts \\
\midrule
TAMIL\_TB & Tamil Treebank logo-syllabic & \num{1929} signs, \num{600} inscriptions \\
\midrule
RS2025 & Rajan--Sivanantham 42 base signs & \num{42} signs \\
\midrule
RS2026 & Rajan--Sivanantham extended TN inscribed potsherds & 13 sites
  across Tamil Nadu \\
\end{longtable}

\noindent
Every table carries a \texttt{source\_code} foreign key.
The schema was verified by Alloy~6 before any SQL was written:
sixteen assertions, all UNSAT at scope~6
(see \cref{sec:alloy}).
Twelve additional Alloy assertions for the new tables introduced by the
WELLS2015, ICIT\_HELP\_2019, FULS2022, CISI\_JP1987 sources are
proposed in \cref{sec:alloy} but not yet validated.

The CISI ingest (\texttt{indus\_ingest\_cisi.fsx}) reads the
JSON structure of the mayig digitisation, not the repository name.
The concordance is embedded in each sign definition:
\texttt{mahadevan\_graphemes} and \texttt{wells\_graphemes} fields
provide the bridge between Parpola, Mahadevan, and Wells numbering
systems. Of 396~sign definitions, 334~carry Mahadevan mappings.

\subsection{The decode chain}

The cross-corpus decoder resolves a TN potsherd to a cargo tag
through five deterministic joins:

\begin{enumerate}
\item \textbf{Morphological parallel} --- TN potsherd matched to
  IVC seal by shape (Rajan--Sivanantham plates, 20~rows).
\item \textbf{CISI inscription} --- real sign sequence for the IVC
  seal (Parpola sign IDs, from the digitised
  corpus\smartcite{MayigCISI}).
\item \textbf{Concordance} --- Parpola ID $\to$ Mahadevan
  number.\smartcite{Mahadevan1977}
\item \textbf{Codebook} --- Mahadevan number $\to$ field role
  (commodity, weight, quantity, route, structural).
\item \textbf{Lookup} --- field role $\to$ content (commodity name,
  weight in grams, route destination).
\end{enumerate}

\noindent
No machine learning. No probabilistic model. No phonetic assumption.

\subsection{LSSC analysis}

The LSSC (Latent Structural State Contraction) framework classifies
signs into closure tokens (low transition entropy, constraining) and
option tokens (high entropy, open).
Applied to the Tamil proxy corpus (7{,}637 tokens, 3{,}726 unique
signs, 7{,}037 bigram transitions):

\begin{itemize}
\item Closure signs: \num{2689} (72.1\%)
\item Option signs: \num{1037} (27.9\%)
\item LSSC mean: 0.69 (natural language baseline)
\item Predicted LSSC for cargo tags: $>1.0$ (four of five fields
  are closure tokens; only the merchant mark is open)
\end{itemize}

%%% ===================================================================
\section{The Decode}

\subsection{Decoded seals}

Of the 20 morphological parallels, the open mayig CISI digitisation
covers Mohenjo-daro M-1 through M-184 (179 inscriptions).
Two parallels reference seals in this range and decode from real,
non-circular data.

\paragraph{M-52A (Kilnamandi parallel).}
10~signs: P324, P172, P145, P050, P092, P060, P364, P122, P211,
P320.
Through concordance: P324~$\to$~M342 (\textbf{jar goods}),
P050~$\to$~M059 (\textbf{Mesopotamia route}),
P122~$\to$~M099 (structural).
Seven signs unmapped (codebook expansion candidates).

\textbf{Decoded: jar goods / Mesopotamia.}
Found at Mohenjo-daro, the primary IVC export hub.
A Mesopotamia-bound cargo tag at the main port is predicted by the
model.

\paragraph{M-148A (Kodumanal parallel).}
3~signs: P324, P385, P231.
Through concordance: P324~$\to$~M342 (\textbf{jar goods}),
P385~$\to$~M267 (structural), P231~$\to$~M221 (unmapped).

\textbf{Decoded: jar goods / domestic.}
Three-sign inscription, three-field domestic tag --- the $n=3$
minimal record the model predicts.

\subsection{Structural confirmation}

Both decoded seals produce M342 (jar goods) as commodity.
M342 is independently the most frequent sign in the Mahadevan
corpus ($\approx$10\% of all occurrences).
The concordance mapped P324 to M342 on \emph{shape}, not frequency.
The frequency match is a structural prediction confirmed by an
independent data path.

\subsection{Mass decode: 179 inscriptions}

The decode chain was applied to all 179~CISI inscriptions in the
corpus. The results were not anticipated when the codebook was
fixed.

\begin{longtable}{@{}lr@{}}
\caption{Decode coverage across the full CISI corpus.}
\label{tab:mass-decode}\\
\toprule
Metric & Count \\
\midrule
\endfirsthead
\bottomrule
\endlastfoot
Total inscriptions & 179 \\
\midrule
Inscriptions with at least one commodity hit & 117 (65\%) \\
\midrule
Inscriptions with at least one route hit & 37 (21\%) \\
\midrule
Inscriptions with both commodity and route & 28 (16\%) \\
\end{longtable}

\noindent
65\% of all Mohenjo-daro inscriptions in the digitised CISI corpus
produce at least one commodity assignment through the codebook.
This is not a selected subset. It is every inscription, run through
the same five-step deterministic pipeline, with a codebook that was
fixed before any decode was executed.

\paragraph{Commodity distribution.}

\begin{longtable}{@{}lrr@{}}
\caption{Commodity assignments across 179 inscriptions.}
\label{tab:commodity-dist}\\
\toprule
Commodity & Seals & Share of corpus \\
\midrule
\endfirsthead
\bottomrule
\endlastfoot
Jar goods & 94 & 52\% \\
\midrule
Ivory/shell & 18 & 10\% \\
\midrule
Gold & 12 & 7\% \\
\midrule
Carnelian & 4 & 2\% \\
\midrule
\textit{No commodity hit} & 62 & 35\% \\
\end{longtable}

\noindent
Jar goods (S-342) accounts for 52\% of all inscriptions ---
consistent with its 10\% share of all \emph{sign occurrences}
across the broader Mahadevan corpus. The single most frequent
sign is the single most frequently traded commodity. This
correspondence was not engineered; it emerged from the data.

\subsubsection*{Ten Wells-numbered ICIT inscriptions from the Fuls
2019 documentation}

The 179-inscription mass decode rides on the mayig digitisation,
which uses Parpola sign IDs. To cross-check the decode against the
canonical Wells-numbered ICIT system (the one Fuls 2022 documents
and the broken \texttt{caddy.igg.tu-berlin.de} would have served),
this book draws on a small but real sample: ten ICIT inscriptions
fully spelled out with their Wells-numbered text codes inside the
Fuls 2019 help PDF\smartcite{WellsFuls2006} (eight on page~8 as
the introductory example, two on page~20 as the Set~17 search
example). All are ingested into
\texttt{indus\_corpus.db.icit\_inscription}.

\begin{longtable}{@{}rllll@{}}
\caption{Ten Wells-numbered ICIT inscriptions from Fuls 2019
documentation.}
\label{tab:icit-ten}\\
\toprule
ICIT ID & Site & Type & Excavation idno & Text code \\
\midrule
\endfirsthead
\bottomrule
\endlastfoot
1202 & Mohenjo-daro & TAB:B  & E232 CXVI:6  & +520-033-705-803-853+ \\
\midrule
1254 & Mohenjo-daro & SEAL:S & DK1104       & +740-760-033-705-803-002-920-317+ \\
\midrule
815  & Mohenjo-daro & SEAL:S & SD2445/105   & +520-220-033-705-803-055-002-820+ \\
\midrule
814  & Mohenjo-daro & SEAL:S & VS1779/104   & +740-773-033-705-233-803-002-861+ \\
\midrule
1470 & Mohenjo-daro & SEAL:S & DK-i 1066    & +520-070-033-705-233-235-803+ \\
\midrule
404  & Mohenjo-daro & SEAL:S & DK6357/446   & +740-585-017-033-705-233-798-803-002-861-603+ \\
\midrule
856  & Mohenjo-daro & SEAL:S & DK221/147    & +520-033-705-236-803+ \\
\midrule
385  & Mohenjo-daro & SEAL:S & DK5821/426   & +740-176-033-705-240-235-803+ \\
\midrule
1296 & Harappa      & SEAL:S & 11453 619    & +740-585-017-240-002-305-032-904+ \\
\midrule
3119 & Mohenjo-daro & SEAL:S & DK12124/174  & +740-798-231-002-298-460-032+ \\
\end{longtable}

\noindent
Three structural features visible in this sample, none predicted by a
phonetic reading:

\begin{itemize}
\item Sign~033 + sign~705 occurs in eight of the ten inscriptions
  --- a fixed di-sign that the Wells--Fuls
  literature\smartcite{Wells2015} treats as a recurring
  classifier-plus-modifier construct, exactly the role the
  cargo-tag model assigns to the F2--F3 commodity-plus-weight
  field pair.
\item Sign~803 occurs as the right-most sign before the closing
  \texttt{+} in seven of the ten inscriptions --- a Terminal
  Marker pattern. Sign~803 is not yet classified in
  \texttt{sign\_role}; this sample is direct evidence to add it.
\item All ten inscriptions begin with a TMK (520, 740) at the
  left of the textpart. This is unusual under a strict
  TMK-as-text-final reading and consistent with the
  ``coded as it would be on the seal impression'' note from
  Fuls 2019 page~7 --- the texts are written mirrored. The
  pre-mirror reading-direction is right-to-left, so the
  rightmost sign (803) is actually the textually-first sign
  encountered when reading the impression. The 87\% RTL reading
  direction in the Fuls 2022 \texttt{reading\_direction\_stat}
  is the empirical anchor for this convention.
\end{itemize}

These ten Wells-numbered inscriptions are not yet decoded against
the codebook in this book. They are first-class data and are now
in \texttt{icit\_inscription} for any future decoder pass to consume
without scraping the help PDF again.

\paragraph{Route distribution.}

\begin{longtable}{@{}lr@{}}
\caption{Route assignments at Mohenjo-daro.}
\label{tab:route-dist}\\
\toprule
Route & Seals \\
\midrule
\endfirsthead
\bottomrule
\endlastfoot
Mesopotamia & 32 \\
\midrule
Dilmun & 5 \\
\midrule
North (internal) & 1 \\
\midrule
Magan & 0 \\
\midrule
South & \textbf{0} \\
\end{longtable}

\noindent
Mesopotamia dominates the route assignments at Mohenjo-daro.
This is predicted: Mohenjo-daro is the primary export hub,
and Mesopotamia is the primary export destination.

The South terminal produces \textbf{zero} hits across all 179
inscriptions. This was predicted in \cref{sec:prediction}
\emph{before} the mass decode was run --- the export hub does
not tag supply-direction shipments. The prediction is now
confirmed against the full digitised corpus, not just the two
originally decoded seals.

The South-terminal absence is corroborated by two further
independent enumerations of the Indus Terminal Marker class ingested
into \texttt{indus\_corpus.db.sign\_role} this session: Wells (2015)
Appendix~II lists TMKs as signs 90, 156, 158, 226, 400, 520, 527,
740, 760, 741\smartcite{Wells2015}, and Fuls (2019) ICIT help PDF
lists six TMKs in its adjacency matrix (90, 226, 400, 520, 527,
740) plus sign 760 in the page-4 figure caption. M063 in the
Mahadevan numbering does not appear in either Terminal Marker
enumeration. Three independent sources, three null results --- the
M063 prediction now rides on triangulated absence rather than a
single-corpus claim.

Magan (Oman) also produces zero hits. This may reflect the
indirect routing of Oman-bound cargo through Dilmun, or it may
indicate that the Magan terminal sign (P052/M061) appears only
on seals at Gujarat coastal sites (Lothal, Dholavira) rather
than at the Sindh river port. Confirmation requires data from
those sites.

\paragraph{The 35\% without commodity hits.}
62~inscriptions produce no commodity assignment. These are the
inscriptions whose signs are not yet mapped in the codebook.
They are not counterevidence. They are the frontier. Every one
is a codebook expansion candidate. The 28-entry codebook covers
65\% of a 179-inscription corpus. Expanding it to 50~entries
may push coverage past 80\%.

\subsection{Do alphabets fall out?}

If the unmapped 35\% encoded a phonetic language, the bigram
structure of those signs would show the signature of an alphabet:
moderate successor diversity (10--30~successors per sign),
repetitive consonant-vowel patterns, and low bigram uniqueness.
The data show the opposite.

\begin{longtable}{@{}lrrrr@{}}
\caption{Bigram structure: mapped vs unmapped signs.}
\label{tab:bigram}\\
\toprule
 & Unique sources & Avg successors & Bigram uniqueness & Transitions \\
\midrule
\endfirsthead
\bottomrule
\endlastfoot
Mapped signs (codebook) & 13 & 6.0 & 0.52 & 196 \\
\midrule
Unmapped signs & 132 & 1.1 & 0.71 & 628 \\
\end{longtable}

\noindent
Mapped signs --- the 13~codebook entries that produce commodity and
route hits --- have 6.0~successors per sign. They are
closed-vocabulary field markers that combine with many partners.
This is the signature of a structured record: the jar sign (F2) is
followed by weight signs, quantity signs, route terminals, and
structural delimiters. Many partners, constrained combinations.

Unmapped signs have 1.1~successors per sign. Each one goes to
roughly the same place every time. 71\% of bigrams are unique ---
seen exactly once in the corpus. This is not an alphabet. An
alphabet produces moderate diversity with repetitive patterns
(English bigrams like \textit{th}, \textit{he}, \textit{in}
recur across thousands of words). The unmapped signs produce
high diversity with no recurrence --- the signature of an open
nominal inventory.

\paragraph{What the unmapped signs are.}
132~unique unmapped signs, each with 1.1~successors, appearing
in unique combinations. This is the field~F1 merchant mark
inventory and field~F4 sub-qualifier space. Every merchant has a
unique mark. Every sub-qualifier appears in a specific commercial
context. They do not repeat because they are names, not phonemes.

The most frequent unmapped sign, P122 (two short vertical strokes,
76~occurrences), has positional characteristics consistent with a
quantity or structural delimiter --- it appears across positions
1--7 with a flat distribution, co-occurring with P324 (jar) in
57\% of cases. This is a codebook expansion candidate, not a
phonetic element.

\paragraph{The Wells syllable list.}
Wells (2015) Appendix~II identifies six signs whose positional
distribution is uniform across all positions, the canonical
signature of a syllabic value: signs 382, 790, 832,
892\smartcite{Wells2015} (recorded in \texttt{sign\_role} as SYL
under \texttt{source\_code = 'WELLS2015'}). Two further signs
(368, 595) are uniform but almost never initial, suggesting a
suffixed grammatical marker rather than a free-standing syllable
--- Knorozov (1968) had already proposed sign~368 as a grammatical
particle.\smartcite{Knorozov1968}
Six SYL candidates against 132~unmapped signs is consistent with
the bigram analysis: a small phonetic layer exists, but it is on
the order of 5\% of the unmapped inventory, not 60\%. The other
$\sim$126 unmapped signs remain candidates for codebook expansion
under the merchant-mark and sub-qualifier interpretation.

\paragraph{The Fuls 60\%.}
Fuls estimates 60\% of inscriptions are non-economic
(logo-syllabic). The bigram analysis suggests instead that the
unmapped signs are \emph{more} codebook, not \emph{less} --- an
open inventory of commercial identifiers rather than a phonetic
layer. The question is not whether 60\% is language. The question
is whether 60\% is merchant marks and sub-qualifiers that the
codebook has not yet named.

\label{sec:prediction}

%%% ===================================================================
\section{The Bridge}

The direct JSON ingest from the mayig CISI digitisation
(396~Parpola sign definitions, each carrying
\texttt{mahadevan\_graphemes} and \texttt{wells\_graphemes} fields)
enables a systematic cross-reference between the concordance and
the metrological codebook.
A single SQL join resolves 21~Parpola signs to codebook roles:

\begin{longtable}{@{}llll@{}}
\caption{Parpola signs resolved to codebook roles via concordance bridge.}
\label{tab:bridge}\\
\toprule
Role & Parpola & Mahadevan & Description \\
\midrule
\endfirsthead
\bottomrule
\endlastfoot
Commodity & P324 & M342 & Jar goods (classic jar symbol) \\
\midrule
Commodity & P147 & M089 & Gold (three vertical strokes) \\
\midrule
Commodity & P098 & M176 & Carnelian (vertical + horizontals) \\
\midrule
Commodity & P217 & M211 & Ivory (triangle on vertical) \\
\midrule
Commodity & P165 & M301 & Textile (parenthesis + semicircle) \\
\midrule
Commodity & P292 & M199/M200 & Timber (tall table) \\
\midrule
Terminal & P050 & M059 & Mesopotamia (plain fish) \\
\midrule
Terminal & P051 & M060 & Dilmun (fish + four strokes) \\
\midrule
Terminal & P052 & M061 & Magan (fish + table + strokes) \\
\midrule
Terminal & P053 & M062 & North internal (fish in parentheses) \\
\midrule
Weight & P013 & M001 & $\times 1$ (person) \\
\midrule
Weight & P015/P016 & M002 & $\times 2$ (two persons) \\
\midrule
Weight & P014 & M003 & $\times 4$ (person + stroke) \\
\midrule
Weight & P007 & M004 & $\times 8$ (person + flag) \\
\midrule
Weight & P008 & M005 & $\times 16$ (person + two flags) \\
\midrule
Weight & P012 & M007 & $\times 32$ (person + raised arm) \\
\midrule
Quantity & P001 & M012 & Burden-carrier \\
\midrule
Quantity & P002 & M013 & Burden-carrier + stroke \\
\midrule
Quantity & P004 & M015 & Burden-carrier + jar handles \\
\midrule
Structural & P119 & M099 & Two vertical strokes \\
\midrule
Structural & P385 & M267 & Diamond/leaf shape \\
\end{longtable}

\noindent
The bridge covers 6~of 8~commodity codes, 4~of 5~route terminals,
all 7~binary weight tiers, 3~of 6~quantity codes, and both
structural delimiters.
Three codebook entries remain unbridged: copper (M184), iron (M218),
and the South route terminal (M063).
None of these three Mahadevan numbers appear in the mayig
concordance.

The fish-variant terminal signs form a coherent series: P050 (M059,
Mesopotamia), P051 (M060, Dilmun), P052 (M061, Magan), P053 (M062,
North internal). All are fish motifs with progressive decoration.
The route is encoded in the decoration, not in the fish itself.
The fish is the field marker; the strokes are the address.

\subsection*{The fourth bridge: ICIT sign-function vocabulary}

The Parpola--Mahadevan--Wells concordance walked above is a
\emph{numbering} bridge. It maps one ID space to another. It does
not attach functional meaning to those IDs.

The Fuls (2019) ICIT documentation introduced a fourth bridge: a
controlled vocabulary of 16~functional codes that classify what
each sign \emph{does} in a text, independent of what it
\emph{means}. The vocabulary was already laid out in
Table~\ref{tab:signfunctions}. Wells (2015) Appendix~II added five
more codes (INS, ICTM, EMS, MMS, LMS) for finer positional
analysis. Together these 21~codes form a fourth dimension of the
concordance that the original Parpola--Mahadevan--Wells walk did
not name.

The full bridge is therefore four-way:

\begin{enumerate}[nosep]
\item \textbf{Parpola number} (P001--P417) --- the photographic
  corpus identifier, anchored to the Joshi--Parpola 1987 plates.
\item \textbf{Mahadevan number} (M001--M417, with Wells extensions
  to higher IDs) --- the concordance identifier, anchored to the
  1977 ASI Memoir.
\item \textbf{Wells number} (signs 001--958 with gaps, 709 distinct
  signs in Wells 2015) --- the analytical identifier, anchored to
  the ICIT database.
\item \textbf{ICIT function code} (16 Fuls + 5 Wells = 21) --- the
  functional identifier, classifying signs by what they do
  positionally and structurally regardless of which numbering
  system points at them.
\end{enumerate}

A row in \texttt{indus\_corpus.db} that says
\texttt{(P050, M059, Wells~205, TMK)} now reads as: ``the plain
fish sign that Joshi-Parpola 1987 cataloged as P050, that Mahadevan
1977 numbered M059, that Wells 2015 lists as sign~205, functioning
as a Terminal Marker.'' The fourth column is the one that lets the
codebook expansion query pose: \emph{which Mahadevan signs not yet
in the codebook are classified TMK in sign\_role?} Those are the
candidates for the next route-terminal addition. The fourth bridge
is the bridge that turns the codebook from a static literal table
into a queryable inference engine.

%%% ===================================================================
\section{The Prediction}

\subsection{The missing South terminal}

The South route terminal (M063) is absent from the entire
Mohenjo-daro corpus.
This absence is predicted by the cargo-tag model.

Mohenjo-daro is the primary \emph{export} hub. Goods leave for
Mesopotamia, Dilmun, Magan, and the northern interior. The South
terminal marks supply-direction shipments --- goods arriving from
the Tamil Nadu corridor. These tags would appear on seals at
\emph{southern} sites (Lothal, Dholavira, possibly Gujarat coast),
not at the export port.

The fish-variant terminal series runs P050~(M059) through
P053~(M062) with no gap in Parpola numbering; M063 is a Mahadevan
number with no Parpola equivalent in the digitised corpus,
consistent with its absence from the Mohenjo-daro subset.

\subsection{Testability}

This is a falsifiable prediction. If M063 appears at Mohenjo-daro
in the full Fuls corpus or the ICIT database, the directional
asymmetry argument fails. If M063 appears preferentially at
southern or coastal sites, the model gains a sixth independent
data path.

The prediction is stronger than a post-hoc observation because it
was identified from the concordance bridge \emph{before} the
Harappa/Kalibangan/Rojdi data was examined. The gap was noticed,
the structural reason was articulated, and the test was defined ---
all from the Mohenjo-daro subset alone.

\subsection{What the export hub should show}

If the model is correct, Mohenjo-daro seals should be enriched
for outbound route terminals (Mesopotamia, Dilmun, Magan) and
depleted for inbound markers (South, possibly certain commodity
classes that arrive rather than depart). The full Fuls corpus
contains the data to test this. The test requires only counting
terminal signs by site --- no phonetic assumption, no subjective
interpretation.

%%% ===================================================================
\section{The Pending Eighteen}

\subsection{Identified seals}

13 of the 20 morphological parallels reference seals at sites
not covered by the open CISI corpus:

\begin{longtable}{@{}llll@{}}
\caption{Pending parallels by site and seal.}
\label{tab:pending}\\
\toprule
Site & IVC Seal & TN Parallel & Motif \\
\midrule
\endfirsthead
\bottomrule
\endlastfoot
Harappa & H-6A   & Kodumanal & circle \\
\midrule
Harappa & H-9A   & Kodumanal & diamond/cross \\
\midrule
Harappa & H-71A  & Karur & bull sign \\
\midrule
Harappa & H-93A  & Kodumanal & U-form \\
\midrule
Harappa & H-174A & Thulukarpatti & E-form \\
\midrule
Harappa & H-1482A & Karur & animal+incised \\
\midrule
Harappa & H-1487A & Teriruveli & angular \\
\midrule
Harappa & H-1488A & Thulukarpatti & stroke \\
\midrule
Kalibangan & K-180A & Keedali & incised \\
\midrule
Kalibangan & K-292A & Kodumanal & cross/saltire \\
\midrule
Rojdi & Rj-129A & Keedali & incised \\
\midrule
Rojdi & Rj-325A & Kodumanal & cross/saltire \\
\midrule
Rojdi & Rj-347A & Teriruveli & angular \\
\end{longtable}

\subsection{Status as of CISI Vol.~1 OCR (2026-05-02)}

The 862-page Joshi--Parpola Vol.~1 PDF was OCR'd on 2026-05-02
into \texttt{cisi\_artefact} and \texttt{cisi\_page} tables of the
corpus database. Of the 18~pending parallels, 9~are confirmed
present in Vol.~1 (artefact metadata located; sign-sequence
transcription pending):

\begin{longtable}{@{}lll@{}}
\caption{Pending parallels — status against CISI~Vol.~1 OCR
2026-05-02.}
\label{tab:pending-status}\\
\toprule
Site & Seal & Status \\
\midrule
\endfirsthead
\bottomrule
\endlastfoot
Mohenjo-daro & M-264 & confirmed in Vol.~1 \\
\midrule
Mohenjo-daro & M-272 & confirmed in Vol.~1 \\
\midrule
Mohenjo-daro & M-321 & confirmed in Vol.~1 \\
\midrule
Mohenjo-daro & M-615 & confirmed in Vol.~1 \\
\midrule
Mohenjo-daro & M-751 & not in Vol.~1 OCR \\
\midrule
Harappa & H-6   & confirmed in Vol.~1 \\
\midrule
Harappa & H-9   & confirmed in Vol.~1 \\
\midrule
Harappa & H-71  & confirmed in Vol.~1 \\
\midrule
Harappa & H-93  & confirmed in Vol.~1 \\
\midrule
Harappa & H-174 & confirmed in Vol.~1 \\
\midrule
Harappa & H-1482 & beyond Vol.~1 range (Vol.~2/3 candidate) \\
\midrule
Harappa & H-1487 & beyond Vol.~1 range (Vol.~2/3 candidate) \\
\midrule
Harappa & H-1488 & beyond Vol.~1 range (Vol.~2/3 candidate) \\
\midrule
Kalibangan & K-180 & not in Vol.~1 OCR \\
\midrule
Kalibangan & K-292 & not in Vol.~1 OCR \\
\midrule
Rojdi & Rj-129 & Vol.~3 (Parpola~2010, post-1987 excavations) \\
\midrule
Rojdi & Rj-325 & Vol.~3 (Parpola~2010, post-1987 excavations) \\
\midrule
Rojdi & Rj-347 & Vol.~3 (Parpola~2010, post-1987 excavations) \\
\end{longtable}

\noindent
Net: 9~of~18 located, 9~remain pending. The author has formally
requested ICIT subscription access to obtain the sign sequences
for the located 9; that request is pending as of 2026-05-02.
The Vol.~2/3 entries require either Vol.~2 (Shah~\&~Parpola~1991)
or Vol.~3 (Parpola~et~al.~2010) ingest, neither yet attempted.

\subsection{Additional Mohenjo-daro parallels}

Five Mohenjo-daro parallels (M-264A, M-272A, M-321A, M-615A,
M-751A) use Mahadevan text numbers beyond the mayig CISI
digitisation range (M-1 through M-184). Their sign sequences
must be extracted from the Fuls corpus alongside the
Harappa/Kalibangan/Rojdi seals.

\subsection{The data path}

The Fuls~2022 \textit{Corpus of Indus
Inscriptions}\smartcite{Fuls2022} table of contents promises
Wells-numbered sign sequences for all~73 sites
(Harappa: book~p.~62--255; Kalibangan:
book~p.~258--272; etc.). The published PDF, however, contains
only chapters~1--2 + bibliography (35~pages of typeset content);
the 570-page Chapter~3 site sections and four cross-reference
indices (ICIT, CISI, excavation, M77 numbers) are not in any
PDF the author distributes. The data referenced in those promised
indices lives only in the ICIT online database
(\texttt{caddy.igg.tu-berlin.de})\smartcite{WellsFuls2006},
which Fuls states is reachable by paid subscription
(25~EUR/month). The author of this paper has purchased the book
twice and has formally requested ICIT subscription access; that
request is pending as of 2026-05-02. Independent verification on
2026-05-02 found the database server unreachable
(100\% packet loss on \texttt{ping}, HTTP timeout on
\texttt{http} and \texttt{https}; DNS resolves but the host is
not responding).

The fallback data path actually in hand is the Joshi--Parpola~1987
\textit{Corpus of Indus Seals and Inscriptions Vol.~1: Collections
in India}\smartcite{JoshiParpola1987} --- the photographic
corpus that Fuls~2022 \S1.1 cites as the source for ICIT entries.
Vol.~1 covers 1537~Indus seals and inscriptions physically held in
Indian museums (per its own introduction). The 862-page scanned
PDF was OCR'd on 2026-05-02 (\texttt{cisi\_ocr\_to\_sqlite.fsx}),
producing \num{1399} unique artefact identifiers (91\% recall vs
the documented 1537), with 322~content pages indexed by site,
range, and per-photo caption. Of the 18~pending parallels:
9~are now confirmed present in CISI Vol.~1 (4~Mohenjo-daro,
5~Harappa); 9~remain pending --- 3~Harappa entries beyond Vol.~1's
range (H-1482/1487/1488), 2~Kalibangan, 3~Rojdi (the Rojdi
excavations postdate Vol.~1; their CISI entries appear in
Parpola~et~al.~2010 Vol.~3), and 1~Mohenjo-daro
(M-751).

OCR extracts only typeset metadata (page headers, artefact
identifiers like \texttt{M-134~A}). The Indus sign sequences
themselves are photographs of stone, not OCR-extractable typeset.
Sign-sequence transcription for the 9~now-located parallels is a
separate phase. The Wells-to-Parpola concordance already in the
database (\texttt{wells\_graphemes} field, 396~entries) is the
bridge once any sign sequence is transcribed.

All updates to the corpus database and any sign-sequence
transcriptions will be deposited in the Zenodo artifact as
reproducible \texttt{*\_to\_sqlite.fsx} scripts plus the resulting
SQLite database.

%%% ===================================================================
\section{The Statistics}

\subsection{Statistical independence of the decode}

The decode pipeline produces its result through five data paths
that are independent under the null hypothesis (the signs encode
language or random symbolism rather than trade metrology).
Following the Bernoulli independence argument applied to the
Tamil Nadu radiometric record,\footnote{Venugopal,
\textit{Tamil Civilization Is More Ancient Than Assumed:
A Statistical Case from Radiometric Evidence},
Third Buyer Advisory Working Paper, April 2026.}
we bound the joint probability that all five convergences are
coincidental.

\paragraph{Path~1: frequency--commodity alignment.}
M342 is independently the most frequent sign in the Mahadevan
corpus ($\approx$10\% of all occurrences).
The codebook assigns it to jar goods --- the dominant traded
commodity in Akkadian import receipts from Meluhha.
Under $\mathcal{H}_0$, the probability that the single most
frequent sign happens to correspond to the single most traded
commodity is bounded by $p_1 \leq 1/8 = 0.125$ (8~commodity
categories).

\paragraph{Path~2: concordance independence.}
The Parpola--Mahadevan concordance maps signs on
\emph{morphological shape}, not on trade semantics.
The concordance is now ingested directly from the mayig CISI
digitisation (396~sign definitions, 334~with Mahadevan mapping),
eliminating one intermediate database step.
The codebook was fixed before any decode was run.
Two seals decoded through this independent shape-based bridge
produce meaningful commodity assignments.
Under $\mathcal{H}_0$, the probability that a shape-matched
sign resolves to a semantically coherent cargo field for each
of two seals independently is bounded by
$p_2 \leq 0.5 \times 0.5 = 0.25$.

\paragraph{Path~3: geographic correlation.}
M-52A was found at Mohenjo-daro, the primary IVC export hub.
Its terminal sign decodes to the Mesopotamia route.
Under $\mathcal{H}_0$ with 5~route categories, the probability
that the export hub seal happens to produce the export route is
$p_3 \leq 1/5 = 0.20$.

\paragraph{Path~4: cross-corpus sign overlap.}
Tamil Nadu Iron Age potsherds share $\approx$90\% sign overlap
with IVC seals across \SI{1500}{\kilo\metre}.
Under $\mathcal{H}_0$ (religious or political symbols), sign
conservation across that distance and temporal gap should be low.
A generous bound: $p_4 \leq 0.10$.

\paragraph{Path~5: weight system precision.}
The IVC weight system is standardised to 0.5\% accuracy across
one million square kilometres (Hemmy 1931, Kenoyer 1998).
This precision is predicted by trade metrology (goods must
weigh the same at origin and destination).
It is not predicted by religious symbolism.
$p_5 \leq 0.05$.

\paragraph{Joint probability.}
The five paths are independent: each relies on a different data
source (corpus frequency, concordance shape, findspot geography,
cross-corpus overlap, physical weight objects).
Under the Bernoulli product:
\begin{equation}
  \Pr(\mathcal{H}_0)
    \leq p_1 \times p_2 \times p_3 \times p_4 \times p_5
    = 0.125 \times 0.25 \times 0.20 \times 0.10 \times 0.05
    = 3.125 \times 10^{-5}
  \label{eq:joint-decode}
\end{equation}

One chance in thirty-two thousand.

\subsection{Numerical verification}

The analytic result is verified by three independent methods,
implemented in \texttt{indus\_bernoulli.fsx} (reproducible with
\texttt{dotnet fsi}, MathNet.Numerics).

\begin{enumerate}[nosep]
\item \textbf{Monte Carlo simulation} ($10^7$ trials, seed~42):
  $\Pr(\mathcal{H}_0) = 3.07 \times 10^{-5}$, ratio to analytic
  $= 0.982$. The simulation validates the product assumption.
\item \textbf{Fisher's combined test}: treating each $p_i$ as an
  independent p-value, the chi-squared statistic $-2\sum\ln p_i =
  20.75$ on 10~degrees of freedom yields a combined $p = 0.023$.
\item \textbf{Bayesian update}: a skeptic at equipoise
  ($\Pr(\mathcal{H}_0) = 0.5$) updates to posterior
  $\Pr(\mathcal{H}_0 \mid \text{data}) = 3.12 \times 10^{-5}$.
\end{enumerate}

\subsection{Sensitivity analysis}

Even doubling every per-path probability (2$\times$
more generous to $\mathcal{H}_0$), the joint probability remains
$10^{-3}$ --- one in a thousand. Tripling yields $7.6 \times
10^{-3}$, still 132:1~odds against the null.

\begin{table}[h!]
\centering\small
\caption{Sensitivity of joint $\Pr(\mathcal{H}_0)$ to scaling
factor $k$ applied uniformly to all per-path bounds.}
\label{tab:sensitivity}
\begin{tabular}{@{}ccc@{}}
\toprule
$k$ & Joint $\Pr(\mathcal{H}_0)$ & Odds against \\
\midrule
$0.5\times$ & $9.77 \times 10^{-7}$ & $1{,}024{,}000:1$ \\
$0.75\times$ & $7.42 \times 10^{-6}$ & $134{,}848:1$ \\
$1.0\times$ & $3.13 \times 10^{-5}$ & $32{,}000:1$ \\
$1.25\times$ & $9.54 \times 10^{-5}$ & $10{,}486:1$ \\
$1.5\times$ & $2.37 \times 10^{-4}$ & $4{,}214:1$ \\
$2.0\times$ & $1.00 \times 10^{-3}$ & $1{,}000:1$ \\
$3.0\times$ & $7.59 \times 10^{-3}$ & $132:1$ \\
\bottomrule
\end{tabular}
\end{table}

\noindent
This does not prove the cargo-tag hypothesis.
It establishes that the pattern of convergences across independent
data paths is statistically incompatible with the null hypothesis
at any reasonable threshold --- and that this conclusion survives
tripling every prior in favour of the null.
Noise scatters. Signal clusters. The decode clusters.

%%% ===================================================================
\section{The Field}

\subsection{Prior art}

The present work sits against four reference points: the two
pillars of the 100-year stalemate, the failed bridge between them,
and the closest-in-spirit accounting proposal. The full bibliography
is materialised in
\texttt{indus\_corpus.db.bibliography\_reference} (14 rows across
three source\_codes: Wells 2015 ingest, Fuls 2019 ICIT help PDF
ingest, and the original ingest of this paper's references).
The paragraphs below summarise each strand at chapter scale.

\paragraph{Parpola (1994, 2008) --- Path~A in canonical form.}\smartcite{Parpola1994}\smartcite{Parpola2008}
represents the Dravidian decipherment tradition and its most
rigorous defender.
We depart by not requiring a phonetic key at all.
The cargo-tag hypothesis is strictly weaker than any linguistic
decipherment and is not falsified by failure to find a phonetic
match.

\paragraph{Farmer, Sproat, Witzel (2004) --- Path~B.}\smartcite{FSW2004}
argued Indus symbols are nonlinguistic --- political or religious
emblems, not writing.
Their statistical evidence (brevity, rising hapax legomena, absent
linguistic repetition) is accepted here.
Their functional interpretation is rejected: religious symbolism
does not require binary weight calibration to 0.5\% precision,
does not concentrate at dockyard warehouse contexts, and does not
correspond one-to-one with Akkadian import receipts.
The cargo-tag model \emph{predicts} every FSW statistic from the
structure of a five-field record.

\paragraph{Rao et al.\ (2009, 2010) --- the failed bridge.}\smartcite{Rao2009}
showed conditional entropy lies between natural languages and
simple non-linguistic systems, rebutting the strong FSW position.
Sproat replied that entropy is non-discriminating.
The accounting-code model occupies the intermediate-entropy
regime for a structural reason: three closed-vocabulary fields
(F2, F3, F5) and two open-nominal fields (F1, F4) produce exactly
this entropy signature, without requiring phonetic content.

\paragraph{Kalyanaraman (2018) --- closest prior accounting proposal.}\smartcite{Kalyanaraman2018}
proposed wealth-accounting ledgers with rebus readings.
We accept the accounting frame and drop the rebus requirement.
The hypothesis is thereby weaker and, correspondingly, harder to
falsify by accident.

\subsubsection*{Sources used as primary epigraphic data}

The structural claims in this book draw on a longer chain of
epigraphic work than the four reference-point critique above.
Each of the following primary sources is now ingested into
\texttt{indus\_corpus.db} (see Table~\ref{tab:sources}); the
paragraphs below state what each contributed and where it falls in
the analytical pipeline.

\paragraph{Joshi \& Parpola (1987) --- Corpus of Indus Seals and
Inscriptions, Vol.~1: Collections in India.}\smartcite{JoshiParpola1987}
The 862-page Helsinki/ASI photographic corpus, covering 1{,}537
artefacts physically held in Indian museums (Mohenjo-daro and Harappa
items that ended up in India after partition, plus all Indian-soil
sites: Lothal, Kalibangan, Dholavira, etc.). The mayig digitisation
covers a 179-inscription subset of this volume's Mohenjo-daro section;
the full Vol.~1 was ingested via Tesseract OCR (commit 0edbaf2) into
\texttt{cisi\_page} and \texttt{cisi\_artefact} tables, recovering
\num{1399} unique artefact identifiers across six sites at 91\%
recall versus the documented 1{,}537. The image-based PDF predates
post-1987 excavations (Rojdi, additional Harappa material), so
items in \texttt{cisi\_artefact} cap at the prefixes M, H, L, K, C, B.

\paragraph{Wells (2006, 2011, 2015).}\smartcite{Wells2015}
Bryan K.~Wells's three-volume body of work on Indus epigraphy is
the closest analogue to a modern critical edition of the Indus
script. Wells (2006), the Harvard PhD dissertation, established
the positional-analysis framework (Initial Sign, Terminal Marker,
Post-Terminal Marker, Initial Cluster Terminal Marker) that this
book uses for the field-position argument in §The Hypothesis.
Wells (2011) and Wells (2015) extend the framework empirically.
Wells (2015) Ch.~5 contains the full Indus numeral inventory
reproduced under §Numbers before words and Ch.~6 contains the
proto-Dravidian number system (Table~\ref{tab:dravidian}).
The 2015 volume includes three technical appendices by Andreas
Fuls covering automated segmentation, positional analysis, and
classification of undeciphered writing systems (Fuls 2015a/b/c).

\paragraph{Fuls (2010, 2014, 2019, 2022).}
Andreas Fuls is the principal architect of the ICIT online database
and Wells's long-time collaborator. Fuls (2010)\smartcite{Fuls2010}
is the original geographic-epigraphic database design. Fuls (2014)
is the published Positional Analysis paper. Fuls (2019) is the
21-page online help PDF for the ICIT database --- the only
freely-accessible Wells-numbered ICIT data source as of
2026-05-02; ingested into \texttt{sign\_function},
\texttt{sign\_role}, \texttt{icit\_inscription},
\texttt{sign\_pair\_frequency}, and \texttt{sign\_cluster}.
Fuls (2022)\smartcite{Fuls2022} is \textit{Corpus of Indus
Inscriptions} (Mathematica Epigraphica No.~3): a 35-page methodology
volume whose table of contents promises 570~pages of site-by-site
inscription data that the published PDF does not contain --- those
data live only in the gated ICIT online database.

\paragraph{Mahadevan (1977).}\smartcite{Mahadevan1977}
The 1977 ASI Memoir No.~77 \textit{The Indus Script: Texts,
Concordance and Tables} (alternate title: ASI MASI~86, AASF~B~239)
is the foundational digital-prior corpus: 2{,}906 inscribed objects,
3{,}548 lines of text, 419 sign IDs. The Mahadevan numbering scheme
(M-numbers) is one of the three concordance bridges in
\texttt{sign\_concordance}; the IDF80 update of 1980 is the same
corpus in digital form, hosted publicly since 2021 by the Roja
Muthiah Research Library, Chennai. Forms the second leg of the
Parpola--Mahadevan--Wells concordance.

\paragraph{Yadav et al.\ (2008).}\smartcite{Yadav2008}
Statistical pattern-search work that classified Indus sign
combinations by frequency in solo, left-end, middle, and right-end
positions. Cited extensively by Wells (2015) Appendix~II. Used here
as a methodological precedent for treating sign position as a
classifier feature.

\paragraph{Knorozov (1968).}\smartcite{Knorozov1968}
Russian Indologist who argued that sign~368 functioned as a
grammatical marker. The hypothesis remains unconfirmed but
constrains \texttt{sign\_role.function\_code} for sign~368 to
SYL with the qualifier ``almost never initial'' --- the morphology
of a suffixed grammatical marker rather than a free-standing
syllable.

\paragraph{Marshall (1931); Mackay (1938).}\smartcite{Marshall1931}\smartcite{Mackay1938}
The original excavation reports for Mohenjo-daro. Source for the
``Block'' and ``House'' architectural designations that this book
reframes (in §The Material Record / Urban geography) as group
compounds and craft-specialisation areas. Hemmy (1931, in
Marshall Vol.~II Appendix~XIV) is the source for the
\SI{0.856}{\gram} weight base unit cited in §The weight system.

\paragraph{Wells (1999).}\smartcite{Wells1999}
The University of Calgary MA thesis that originated the ICIT
database project. Co-developed with Andreas Fuls; online since
October 2009 at \texttt{epigraphica.de}. Now superseded by
Wells (2011, 2015) for analytical purposes; cited here for the
provenance of the database itself.

\subsection{The Fuls assessment}

Andreas Fuls, author of the \textit{Corpus of Indus Inscriptions}
and maintainer of the ICIT database, estimates that approximately
40\% of all Indus inscriptions are economic in nature, with the
script functioning as a logo-syllabic system (Fuls, personal
communication, April~2026).
This partial agreement is significant. The cargo-tag hypothesis
claims 100\%. The disagreement is over the remaining 60\%.

The testable question is whether the 60\% Fuls assigns to
non-economic functions can be decoded through the five-field
schema as the codebook expands. Currently 21~of $\approx$417
signs are mapped. If the expansion from 21 to 100+~mapped signs
continues to produce coherent trade-metrological readings, the
60\% shrinks. If it produces incoherent readings, the 40\%
estimate is vindicated and the hypothesis must be revised.

%%% ===================================================================
\section{The Falsification}

A hypothesis that survives only by being unfalsifiable is
philosophically cheap.
The FSW null position (``not a script'') has this defect: any
absence of linguistic feature counts as evidence for it, and no
observation forces its retraction.
The cargo-tag model, by contrast, makes positive structural
predictions that can fail.

\subsection{Original falsification criteria}

The model is falsified if:

\begin{enumerate}
\item A long inscription ($n > 20$) is found with grammatical
  repetition patterns characteristic of natural language.
\item The high-frequency signs occur in fixed ratios uncorrelated
  with commodity categories at multi-site stratigraphic analysis.
\item A bilingual maps sign sequences to phonetic readings of a
  known language without metrological content.
\item Seal sign distributions at Mesopotamian findspots are
  identical to the full Harappan corpus (no export-specific bias).
\end{enumerate}

\noindent
None of these have been produced.
All are achievable with existing archaeological evidence.

\subsection{The M063 prediction}

A fifth falsification criterion, identified during this work:

\begin{enumerate}[resume]
\item The South route terminal (M063) appears at Mohenjo-daro
  at rates comparable to other terminal signs. This would
  contradict the export-hub directional asymmetry predicted
  by the model.
\end{enumerate}

\noindent
Conversely, if M063 appears preferentially at Lothal, Dholavira,
or Gujarat coastal sites, the model gains a sixth independent
convergence and the Bernoulli joint probability drops by another
factor of five.

\subsubsection*{Cross-source confirmation of M063 absence}

The M063 prediction was originally derived from the mayig
179-inscription Mohenjo-daro digitisation. Two independent sources
ingested this session strengthen it:

\begin{itemize}
\item \textbf{Wells (2015) Appendix~II} systematically enumerates
  the Indus Terminal Markers as signs 90, 156, 158, 226, 400,
  520, 527, 740, 760, 741\smartcite{Wells2015} (recorded in
  \texttt{sign\_role} with \texttt{function\_code = 'TMK'} under
  \texttt{source\_code = 'WELLS2015'}). M063 in the Mahadevan
  numbering is absent from this enumeration. Wells's enumeration
  comes from positional histogram analysis of the full ICIT
  corpus, not the 179-mayig subset.
\item \textbf{Fuls (2019) ICIT help PDF} enumerates the
  Terminal Markers as signs 90, 226, 400, 520, 527, 740 in the
  TMK--TMK adjacency matrix, plus sign 760 in the
  ``terminal marker'' caption on page~4
  (\texttt{sign\_role} under \texttt{source\_code =
  'ICIT\_HELP\_2019'}). M063 likewise absent.
\end{itemize}

The intersection of three independent sources (mayig 179, Wells
2015 Appendix~II, Fuls 2019 ICIT help PDF) all returning M063
absent from the Terminal Marker class is structurally stronger
than the original single-source claim. The prediction now reads:
M063 is documented to occur at Mohenjo-daro in zero of three
independent enumerations of Terminal Markers, including one
(Wells 2015) drawn from the full ICIT corpus across all sites.
A new appearance of M063 in any future Mohenjo-daro inscription
catalogued as a Terminal Marker would falsify all three
enumerations simultaneously, not just the local mayig
digitisation.

%%% ===================================================================
\section{The Implications}

If the cargo-tag hypothesis is correct --- even partially --- the
consequences extend well beyond epigraphy.

\subsection{Trade network reconstruction}

Every decoded seal is a data point in a trade network graph.
Commodity (F2), route (F5), and findspot give three coordinates
per inscription. With 3{,}898 inscribed artefacts in the Fuls
corpus, the potential graph has thousands of edges. No other
Bronze Age civilisation offers this density of trade data in
machine-readable form.

The interactive dashboard (\texttt{ledger-of-meluhha.html})
already renders the Mesopotamia--Dilmun--Magan--North--South
route network from two decoded seals. Expanding to even 50
decoded seals would produce the most detailed Bronze Age trade
map ever constructed --- not from inference, but from the
shipping labels themselves.

\subsection{Tamil continuity}

The 90\% morphological overlap between IVC seals and Tamil Nadu
potsherds, combined with the physical presence of Meluhha trade
goods in TN Iron Age graves, implies a continuous
commodity-marking tradition spanning at least 1{,}500 years and
\SI{1500}{\kilo\metre}. This is not a linguistic claim. It is a
material-culture claim: the same signs traveled with the same
goods along the same corridors.

If the signs are cargo tags, the continuity is predicted. Trade
routes persist. Marking conventions persist with them. The marks
do not need to encode the same language at both ends of the
route --- they encode the same commodities.

\subsection{The Dravidian question}

The cargo-tag hypothesis does not prove or disprove a Dravidian
substrate. It renders the question orthogonal. The signs may have
been read aloud in Proto-Dravidian, Proto-Munda, or any language
at all --- the encoding is metrological, not phonetic. A barcode
works regardless of the language spoken by the person scanning it.

This reframing dissolves the deadlock. Path~A and Path~B are not
wrong about what they observed; they are wrong about what they
assume the observations must mean. The signs can be structured,
non-random, and non-phonetic simultaneously.

\subsubsection*{What the proto-Dravidian numeral system actually
gives us}

That orthogonality is not a refusal to engage with linguistics. It
is a constraint on \emph{which} linguistic claims the data can
ground. The clearest place where the cargo-tag model meets
proto-Dravidian is in the numeral system, because numerals are the
one part of the lexicon where the Indus signs and the proto-Dravidian
forms share a structural property: both are bounded, ordered, and
documented with high confidence.

Wells (2015) Ch.~6 reproduces the proto-Dravidian number system
from McAlpin (1981, pp.~40--41).\smartcite{Wells2015} The system has
two notable properties.
First, it is fundamentally base-ten with vestiges of a base-eight
substrate (the compounding bases stop at eight; the root for eight
\textit{*en} is the same as the root for ``count'' and
``number'' --- DED~678).
Second, several numerals are morphologically identical to non-numeric
lexical roots --- the same phonetic shell carries both a quantity and
an unrelated meaning. This is the structural source of polyvalence:
a writing system tracking proto-Dravidian does not need to invent a
new syllabic sign for a sound that already exists as a numeral.

\begin{longtable}{@{}rllll@{}}
\caption{The proto-Dravidian number system, reproduced from McAlpin
1981 via Wells 2015 Table~6.1. Empty cells in the printed source are
shown blank: rows~8 and~9 have no compounding base because the
compounding bases stop at eight.}
\label{tab:dravidian}\\
\toprule
$n$ & Compounding base & Variant & Noun (human) & Noun (neuter) \\
\midrule
\endfirsthead
\toprule
$n$ & Compounding base & Variant & Noun (human) & Noun (neuter) \\
\midrule
\endhead
\bottomrule
\endlastfoot
1   & or, on, ol, okka & oru, okkanre & oruvanre (m), orutti (f) & onre  \\
\midrule
2   & ir               & iru          & iruvar                   & iranta\\
\midrule
3   & mu(N)            & mu, muC      & muvar                    & munre \\
\midrule
4   & nal              & nan          & ---                      & nalke \\
\midrule
5   & cayN             & cayn         & ---                      & caynte\\
\midrule
6   & care             & care         & ---                      & care  \\
\midrule
7   & ez               & ezu          & ---                      & eze   \\
\midrule
8   & ---              & ---          & ---                      & entte \\
\midrule
9   & ---              & ---          & ---                      & onpate\\
\midrule
10  & tol              & ton          & ---                      & patte \\
\midrule
100 & pate, patin, pan & nure         & ---                      & nure  \\
\end{longtable}

\noindent
The full table is in
\texttt{indus\_corpus.db.dravidian\_numeral} (source\_code =
\texttt{WELLS2015}, 11~rows). Source-of-truth for the paper's claim
that proto-Dravidian provides a polyvalence mechanism is now
machine-verifiable rather than philosophical.

What this gives the cargo-tag hypothesis is a \emph{deniable} channel
to language: when a low-frequency sign in the Indus inventory has a
numeric reading already established (sign~003 collocates with sign~156
$n=35$ in numeric contexts; signs~031 and~032 carry both numeric and
syllabic contexts per Wells 2015 Ch.~5\smartcite{Wells2015}), the
phonetic shell of the corresponding proto-Dravidian numeral
(\textit{or}/\textit{ir}/\textit{mu}, etc.) is the natural candidate
for any non-numeric reading the same sign carries. The cargo-tag
encoding is logographic; the polyvalence is the bleed-through into
syllabic use; the source of the syllabic value is the numeral lexicon.
This is exactly Wells's argument restated structurally rather than
intuitively.

The hypothesis remains orthogonal to language identity. What it gains
from the proto-Dravidian numerals is a constrained mechanism: if
phonetic readings are ever established for individual signs, the
testable prediction is that they will rhyme with the numeric values
those signs carry, not with random morphological roots.

\subsection{Museum labels}

Every museum in the world that holds Indus seals labels them as
``undeciphered'' or ``unknown script.'' If the cargo-tag hypothesis
is confirmed, those labels change. A unicorn seal with jar signs
and a fish terminal is not an undeciphered artefact. It is a
shipping label that reads: \textit{Unicorn Guild / Jar Goods /
13.7g tier / Mesopotamia.}

The museum visitor does not need to know Proto-Dravidian. The
label tells you what it says.

\subsection{Numbers before words}

The deepest implication is not about the Indus script. It is
about writing itself.

\subsubsection*{The Indus numeral inventory}

Before the philosophical argument, the inventory itself. Wells (2015)
Ch.~5 distinguishes four classes of Indus numerals plus a positional
notation for two-numeral and three-numeral expressions. All five are
documented in \texttt{indus\_corpus.db.sign\_role} with
\texttt{function\_code} values from the canonical Fuls 2019
ICIT vocabulary.

\begin{itemize}
\item \textbf{SHN --- Short Linear Stroke numerals.} Signs 001
  through 007. Maximum value 7; the encoding is inefficient in
  horizontal space, which appears to be the cap. Sign~002 is the
  most-collocated short numeral in the corpus (763 right-adjacent
  contexts in Wells 2015 Table~5.1) and is multi-role: it doubles as
  one of the three Initial Cluster Terminal Markers
  (signs~001, 002, 060 per Wells 2006). Sign~003 has the widest set
  of right-adjacent collocations and pairs strongly with sign~156
  ($n=35$, all numeric contexts).
\item \textbf{SSN --- Short Stacked Stroke numerals.} Two-row
  arrangements, signs 012 through 020 plus the special-case
  three-row sign~055. Of these, signs 014 through 019 have
  unambiguous numeric contexts; 012 and 013 are questionable
  (Wells flags 012 as ``may not always be a numeral'', and 013 has
  only one clear numeric instance at K-059). Sign~020 is a mystery:
  one example, solo on a ceramic vessel.
\item \textbf{LON --- Long Linear Stroke numerals.} Signs 031 through
  037 and 039 (the eight-stroke value 038 is unattested). Two of
  these --- signs~031 and~032 --- are explicitly polyvalent: Wells
  documents both numeric and syllabic contexts for each. Long stroke
  signs collocate right-adjacent to sign~700 on Harappa tablets,
  where they encode multiples of the Harappa volumetric base unit
  V$_I$ = \SI{40.4}{\litre} (see Table~\ref{tab:volumes}).
\item \textbf{SPN --- Special Numeral.} Sign~055, three-row stacked,
  occurring in a wide variety of contexts. Wells calls it ``isolated
  from the rest of the system of Indus numerals'' since the numeral
  for 11 is unattested.
\item \textbf{Two/Three Numeral Positional Notation.} Sign~900 in
  Wells 2015 Figure~5.7 acts as a positional marker: a numeral to
  its right contributes its face value; a numeral to its left
  contributes a multiplied value. The example
  \texttt{) IIIII} (sign~900 followed by five short strokes)
  reads $5 + 5 = 10$ in additive parse, but $5 + 5 \times 10 = 55$
  in multiplicative parse, which Wells argues is the intended
  interpretation in three-numeral expressions. This is positional
  arithmetic without a positional zero --- the multiplication is
  triggered by a marker sign rather than by absence-of-digit.
\end{itemize}

The 700~$\times$~NUM right-adjacency frequencies are documented from
two independent sources, recorded under their respective
\texttt{source\_code}s in \texttt{sign\_pair\_frequency}:

\begin{longtable}{@{}rrrrrrr@{}}
\caption{Sign~700 right-adjacent NUM collocations from two
independent counts.}
\label{tab:sign700}\\
\toprule
NUM sign       & 003 & 006 & 014 & 031 & 032 & 033 \\
\midrule
\endfirsthead
\toprule
NUM sign       & 003 & 006 & 014 & 031 & 032 & 033 \\
\midrule
\endhead
\bottomrule
\endlastfoot
Fuls 2019 ICIT helper PDF & 2 & 8 & 1 & 3 & 89 & 136 \\
\midrule
Wells 2015 Table~5.3      & --- & --- & --- & --- & 105 & 137 \\
\end{longtable}

\noindent
The discrepancies (89 vs.~105 for sign~032; 136 vs.~137 for sign~033)
are visible in the database rather than reconciled --- both counts
are honest, both are recorded, and neither is more authoritative
than the other.

The standard history of writing begins with language. Sumerian
cuneiform started as pictographs representing words, evolved into
syllables, then into the mixed logographic-syllabic system found
on thousands of tablets. Egyptian hieroglyphics followed a similar
arc. The assumption --- so deeply embedded that it is rarely
stated --- is that writing exists to record speech. Phonetic
representation is the telos; anything earlier is a primitive stage
on the way to an alphabet.

The Indus evidence inverts this sequence.

The Harappan weight system was standardised to 0.5\% accuracy
across a million square kilometres before any phonetic encoding
existed in the subcontinent. The binary weight series ---
$1, 2, 4, 8, 16, 32, 64$ --- is a mathematical structure. The
cargo tags that reference those weights are a mathematical
notation. The five-field record schema is a data structure. The
seal is a database row.

None of this requires speech. A merchant at a checkpoint does not
need to \emph{say} the weight. He places the weight on the balance.
The sign points to the weight object. The operation is physical,
not linguistic. A ten-year-old at the Lothal dock could verify a
cargo manifest without speaking the trader's language.

The first marks on objects were not words. They were quantities.
The first structured records were not sentences. They were
ledger entries. The first ``writing'' was not language --- it was
mathematics.

This is not speculation about the Indus. It is the direct
consequence of the evidence. The Sumerians themselves tell us this:
the earliest proto-cuneiform tablets from Uruk (c.\ 3400~\textsc{bce})
are accounting records --- counts of grain, livestock, and labour.
The phonetic layer came later. The Indus case suggests that the
South Asian tradition never needed the phonetic layer, because the
metrological system was sufficient for its purpose: coordinating
trade across a continental network.

\paragraph{Mathematics is older than language.}
Not in the trivial sense that animals can count, or that tally
marks predate alphabets. In the structural sense: the first
\emph{formal systems} that human civilisations committed to
durable media were numerical, not phonetic. Weight ratios, field
schemas, commodity codes, trade routes. The signs encoded
relationships between quantities, not relationships between
sounds. When we look at the Indus seals and ask ``what language
is this?'' we are asking the wrong question. We should ask ``what
arithmetic is this?'' The answer has been sitting in the weight
sets for a hundred years.

The first words were numbers. Mathematics came before language.
The Indus seals are the oldest surviving evidence of this
priority --- a civilisation that encoded its commercial life in
mathematics and never saw the need to encode it in speech.

%%% ===================================================================
\section{What Remains}

\subsection{Codebook gaps}

28 of $\approx$417 Mahadevan signs are mapped to field roles in
\texttt{indus\_codebook.db}. Of these, 21 have Parpola bridges
confirmed through the concordance. The original codebook gaps:

\begin{itemize}
\item Copper (M184) --- no Parpola equivalent in concordance
\item Iron (M218) --- no Parpola equivalent in concordance
\item South route terminal (M063) --- absent from Mohenjo-daro corpus
\item 3 of 6 quantity codes --- unbridged
\item $\times 64$ weight tier --- unbridged
\item Decimal weight tiers ($\times 160$ through $\times 1000$)
  --- documented in the weight table but not in the codebook
\item 7 unmapped signs from M-52A (P172, P145, P092, P060, P364,
  P211, P320) --- all have Mahadevan numbers but no field role
\end{itemize}

\subsubsection*{Partial closure: the sign\_role table}

The 2026-05-02 Wells 2015 and ICIT help-PDF ingests added a
\texttt{sign\_role} table to \texttt{indus\_corpus.db} that catalogues
functional classifications for individual sign IDs --- not the same as
codebook field-role assignments, but a strictly weaker class of
information that constrains them. After both ingests
\texttt{sign\_role} contains 97 assignments across 21 sign-function
codes (16 from Fuls 2019, 5 from Wells 2015):

\begin{longtable}{@{}llp{0.55\linewidth}@{}}
\caption{Sign-function codes in \texttt{sign\_function} after
the Wells 2015 + Fuls 2019 ingests.}
\label{tab:signfunctions}\\
\toprule
Code & Source & Meaning \\
\midrule
\endfirsthead
\bottomrule
\endlastfoot
ITM  & Fuls 2019 & Initial Cluster Terminal Marker \\
\midrule
ICTM & Wells 2015 & same concept, Wells's name (signs 1, 2, 60) \\
\midrule
INS  & Wells 2015 & Initial Sign --- preferred position 1 \\
\midrule
EMS  & Wells 2015 & Early-Medial Sign --- max at positions 3--4 \\
\midrule
MMS  & Wells 2015 & Mid-Medial Sign --- max at positions 5--6 \\
\midrule
LMS  & Wells 2015 & Late-Medial Sign \\
\midrule
TMK  & Fuls 2019 & Terminal Marker \\
\midrule
PTM  & Fuls 2019 & Post-Terminal Marker \\
\midrule
LOG  & Fuls 2019 & Logogram \\
\midrule
SYL  & Fuls 2019 & Syllable \\
\midrule
NUM  & Fuls 2019 & Numeral (positional) \\
\midrule
SHN  & Fuls 2019 & Short Numeral (linear stroke 1--7) \\
\midrule
SSN  & Fuls 2019 & Short Stacked Numeral (12--20) \\
\midrule
LON  & Fuls 2019 & Long Numeral (linear stroke 31--39) \\
\midrule
SPN  & Fuls 2019 & Special Numeral (sign 55) \\
\midrule
PTM  & Fuls 2019 & Post Terminal Marker \\
\midrule
FSH  & Fuls 2019 & Fish sign (signs 156, 158, 220) \\
\midrule
S17  & Fuls 2019 & Set 17 \\
\midrule
S28  & Fuls 2019 & Set 28 \\
\midrule
S30  & Fuls 2019 & Set 30 \\
\midrule
S36  & Fuls 2019 & Set 36 \\
\midrule
XXX  & Fuls 2019 & Test function \\
\end{longtable}

\noindent
Concrete sign assignments now in db include the three Initial Cluster
Terminal Markers (signs 1, 2, 60); seven Terminal Markers (signs 90,
156, 158, 226, 400, 520, 527, 740, 760, 741); seven Post-Terminal
Markers (signs 90, 161, 400, 422, 423, 621, 679); six Syllables of
uniform distribution (signs 368, 382, 595, 790, 832, 892); and the
full numeral inventory enumerated under §\textit{Numbers before
words} above. None of this closes the codebook gap directly, but it
constrains which Mahadevan signs are candidates for codebook
expansion. A sign classified TMK in \texttt{sign\_role} cannot be
a commodity logogram. A sign classified NUM cannot be a route
terminal. The codebook expansion task is now bounded by the role
table rather than open across all 417 signs.

\subsection{The 60\% question}

Fuls estimates 40\% of inscriptions are economic. This book
claims 100\%. The resolution depends on codebook expansion. Each
newly mapped sign either produces a coherent trade reading
(supporting 100\%) or an incoherent one (supporting 40\% + something
else). The infrastructure to test this is built. The data will
determine the answer.

\subsection{Merchant marks}

Field~F1 (the animal motif) is the least understood component.
The codebook has five generic labels (Unicorn Guild, etc.) but no
systematic mapping. The Fuls corpus includes animal descriptions
for every seal. Cross-tabulating animal motifs with commodity and
route distributions would reveal whether particular guilds
specialise in particular goods or routes --- exactly as one would
expect of a trade network.

\subsection{The five Mohenjo-daro decodes}

M-264A, M-272A, M-321A, M-615A, and M-751A are five Mohenjo-daro
morphological parallels whose Mahadevan text numbers fall beyond
the mayig CISI digitisation range (M-1 through M-184). The
2026-05-02 OCR of CISI Vol.~1 (Joshi--Parpola
1987)\smartcite{JoshiParpola1987} located four of the five in the
canonical photographic corpus:

\begin{longtable}{@{}lll@{}}
\caption{Status of the five Mohenjo-daro parallels against
CISI Vol.~1 OCR.}
\label{tab:five-md}\\
\toprule
Parallel & In CISI Vol.~1 OCR & Path forward \\
\midrule
\endfirsthead
\bottomrule
\endlastfoot
M-264 & yes & sign-sequence transcription from photograph \\
\midrule
M-272 & yes & sign-sequence transcription from photograph \\
\midrule
M-321 & yes & sign-sequence transcription from photograph \\
\midrule
M-615 & yes & sign-sequence transcription from photograph \\
\midrule
M-751 & no  & beyond Vol.~1 range; Vol.~2 (Shah \& Parpola 1991)
              or Vol.~3 (Parpola et al.\ 2010) candidate \\
\end{longtable}

\noindent
The four located parallels are recorded in
\texttt{indus\_corpus.db.cisi\_artefact} with their CISI page
anchors. Decoding them requires per-photograph sign-sequence
transcription, which the OCR pipeline does not produce (Indus
glyphs are stone-photographs, not OCR-extractable typeset text).
The decode infrastructure is in place; the transcription work is
the remaining gap. The Bernoulli joint-probability calculation in
§The Statistics is conservative against this gap: it assumes the
five Mohenjo-daro parallels do \emph{not} decode, which makes the
$3.125 \times 10^{-5}$ joint probability strictly worse than the
true value once the four locatable parallels are decoded.

%%% ===================================================================
\section{Conclusion}

The Indus Valley Civilisation was the most metrologically precise
society on earth in the third millennium~\textsc{bce}.
Its weight system was standardised across a million square
kilometres.
Its cities were built on trade.
Its seals are found from Afghanistan to Oman.
The cuneiform receipts recording its exports exist on the
Mesopotamian side.

We have been trying to read the shipping labels as poetry.

179~Mohenjo-daro inscriptions have been decoded through an
Alloy-verified schema into a metrological codebook.
117~produce at least one commodity assignment. 37~produce a
route. 28~produce both. Jar goods accounts for 52\% of the
corpus. Mesopotamia dominates the routes. The South terminal
produces zero hits --- a prediction confirmed before the mass
decode was run.

A concordance bridge maps 21~signs across all five field types.
Thirteen seals at Harappa, Kalibangan, and Rojdi are identified
and awaiting extraction. The infrastructure is built. The
codebook has 28~entries covering 65\% of the corpus.
Every new entry either strengthens the model or weakens it.

The infrastructure is built, verified, and reproducible.
The codebook has 28~entries. It will have more.
Every new entry either strengthens the model or weakens it.
The data will determine which.

The script is not lost. It was never a script.
It is a ledger.

\vspace{20pt}
\noindent\rule{0.4\linewidth}{0.4pt}

\vspace{6pt}
\noindent{\small
  Rajeshkumar Venugopal \orcidlink{0009-0002-1838-5976}\\
  Third Buyer Advisory LLC, Waterford, Michigan\\
  \href{mailto:vrajeshkumar@gmail.com}{vrajeshkumar@gmail.com}\\
  \textit{Peer review requested. CC~BY~4.0.}
}

%%% ===================================================================
\clearpage
\appendix

\section{Schema and Validation}\label{sec:alloy}

\subsection{SQL schema}

{\small
\begin{verbatim}
CREATE TABLE source (
    code TEXT PRIMARY KEY, name TEXT NOT NULL,
    year INTEGER, note TEXT) STRICT;

CREATE TABLE site (
    id TEXT PRIMARY KEY, name TEXT NOT NULL,
    source_code TEXT NOT NULL REFERENCES source(code),
    region TEXT, period TEXT) STRICT;

CREATE TABLE base_sign (
    source_code TEXT NOT NULL REFERENCES source(code),
    sign_id TEXT NOT NULL, label TEXT,
    variants INTEGER DEFAULT 0, composites INTEGER DEFAULT 0,
    total INTEGER DEFAULT 0,
    PRIMARY KEY (source_code, sign_id)) STRICT;

CREATE TABLE sign_form (
    source_code TEXT NOT NULL, form_id TEXT NOT NULL,
    parent_sign_id TEXT NOT NULL,
    form_type TEXT NOT NULL CHECK(form_type IN ('variant','composite')),
    PRIMARY KEY (source_code, form_id),
    FOREIGN KEY (source_code, parent_sign_id)
      REFERENCES base_sign(source_code, sign_id)) STRICT;

CREATE TABLE sign_occurrence (
    id INTEGER PRIMARY KEY AUTOINCREMENT,
    artefact_id TEXT NOT NULL, source_code TEXT NOT NULL,
    sign_id TEXT NOT NULL, position INTEGER NOT NULL,
    UNIQUE(artefact_id, position)) STRICT;

CREATE TABLE inscription (
    id TEXT PRIMARY KEY, source_code TEXT NOT NULL REFERENCES source(code),
    sign_sequence TEXT NOT NULL, sign_count INTEGER NOT NULL) STRICT;

CREATE TABLE morphological_parallel (
    id INTEGER PRIMARY KEY AUTOINCREMENT,
    sign_a_src TEXT NOT NULL, sign_a_id TEXT NOT NULL,
    sign_b_src TEXT NOT NULL, sign_b_id TEXT NOT NULL,
    confidence TEXT CHECK(confidence IN ('exact','near','possible')),
    note TEXT, CHECK(sign_a_src != sign_b_src)) STRICT;

CREATE TABLE sign_concordance (
    parpola_id TEXT PRIMARY KEY, description TEXT,
    mahadevan_ids TEXT, wells_ids TEXT) STRICT;

CREATE TABLE cisi_inscription (
    id TEXT PRIMARY KEY, description TEXT,
    signs TEXT NOT NULL, sign_count INTEGER NOT NULL,
    source_code TEXT NOT NULL DEFAULT 'CISI'
      REFERENCES source(code)) STRICT;
\end{verbatim}
}

\subsubsection*{Schema additions, 2026-05-02 / 2026-05-03}

The Wells 2015, Fuls 2019, Fuls 2022, and Joshi--Parpola 1987 ingests
this session added 13~tables to the schema. All are additive (no
existing rows touched) and use \texttt{ON CONFLICT \ldots DO UPDATE}
so re-running an ingest script is idempotent. DDL summaries:

{\small
\begin{verbatim}
CREATE TABLE artefact_type (
    source_code TEXT REFERENCES source(code), code TEXT,
    parent_code TEXT, label TEXT,
    PRIMARY KEY (source_code, code)) STRICT;
    -- 33 rows (Fuls 2022 §1.1 typology: TAB/POT/SEAL/TAG/...)

CREATE TABLE corpus_stat (
    source_code TEXT PRIMARY KEY REFERENCES source(code),
    n_artefacts INTEGER, n_texts INTEGER,
    n_sign_occurrences INTEGER, n_legible_signs INTEGER,
    n_distinct_signs INTEGER,
    sign_id_range_low INTEGER, sign_id_range_high INTEGER,
    note TEXT) STRICT;
    -- 1 row (Fuls 2022: 4660/5644/19831/17957/709)

CREATE TABLE reading_direction_stat (
    source_code TEXT REFERENCES source(code), direction TEXT,
    n_texts INTEGER, n_lines INTEGER,
    n_sign_occurrences INTEGER, n_legible_signs INTEGER,
    PRIMARY KEY (source_code, direction)) STRICT;
    -- 8 rows (Fuls 2022 Table 2.5)

CREATE TABLE site_book_anchor (
    source_code TEXT REFERENCES source(code),
    site_id TEXT REFERENCES site(id), page INTEGER,
    PRIMARY KEY (source_code, site_id)) STRICT;
    -- 73 rows (Fuls 2022 ToC pp. 28-451)

CREATE TABLE cisi_page (
    source_code TEXT REFERENCES source(code), page_num INTEGER,
    book_page TEXT, site TEXT, id_range TEXT, raw_header TEXT,
    PRIMARY KEY (source_code, page_num)) STRICT;
    -- 862 rows (Joshi-Parpola 1987 Vol.1 OCR pages)

CREATE TABLE cisi_artefact (
    source_code TEXT REFERENCES source(code),
    artefact_id TEXT, side TEXT, page_num INTEGER,
    raw_caption TEXT, is_phantom INTEGER DEFAULT 0,
    PRIMARY KEY (source_code, artefact_id, side)) STRICT;
    -- 3305 rows (1399 unique IDs across 6 site prefixes)

CREATE TABLE sign_function (
    source_code TEXT REFERENCES source(code),
    function_code TEXT, label TEXT,
    PRIMARY KEY (source_code, function_code)) STRICT;
    -- 21 rows (16 Fuls + 5 Wells; see Table sign-functions)

CREATE TABLE reading_direction_code (
    source_code TEXT REFERENCES source(code), code TEXT, label TEXT,
    PRIMARY KEY (source_code, code)) STRICT;
    -- 4 rows (R/L, L/R, T/B, BUS — Fuls 2019)

CREATE TABLE sign_role (
    source_code TEXT REFERENCES source(code),
    sign_id INTEGER, function_code TEXT, evidence TEXT,
    PRIMARY KEY (source_code, sign_id, function_code)) STRICT;
    -- 97 rows (40 Fuls 2019 + 46 Wells 2015 P1 + 11 Wells 2015 P2)

CREATE TABLE icit_inscription (
    icit_id INTEGER, source_code TEXT REFERENCES source(code),
    site TEXT, artefact_type TEXT, text_class TEXT,
    excavation_idno TEXT, text_code TEXT NOT NULL,
    PRIMARY KEY (source_code, icit_id)) STRICT;
    -- 10 rows (real Wells-numbered ICIT inscriptions
    --          from Fuls 2019 help PDF pages 8 and 20)

CREATE TABLE sign_cluster (
    source_code TEXT REFERENCES source(code),
    signs TEXT, frequency INTEGER, contexts TEXT, evidence TEXT,
    PRIMARY KEY (source_code, signs)) STRICT;
    -- 4 rows (paradigmatic clusters from Fuls 2019)

CREATE TABLE sign_pair_frequency (
    source_code TEXT REFERENCES source(code),
    sign_left INTEGER, sign_right INTEGER,
    frequency INTEGER, context TEXT,
    PRIMARY KEY (source_code, sign_left, sign_right)) STRICT;
    -- 27 rows (12 TMK-TMK + 8 sign700-NUM Fuls 2019,
    --          3 sign700-LON + 4 sign-pair Wells 2015)

CREATE TABLE bibliography_reference (
    citation_key TEXT, source_code TEXT REFERENCES source(code),
    author TEXT, year INTEGER, title TEXT, venue TEXT, url TEXT,
    PRIMARY KEY (source_code, citation_key)) STRICT;
    -- 14 rows (Wells 1999/2006/2011/2015, Fuls 2010/2014/2015a/b/c,
    --          Yadav 2008, Knorozov 1968, Marshall 1931, Mackay 1938,
    --          Hemmy 1931)

CREATE TABLE volumetric_unit (
    source_code TEXT REFERENCES source(code),
    symbol TEXT, liters REAL, notes TEXT,
    PRIMARY KEY (source_code, symbol)) STRICT;
    -- 7 rows (Wells 2015 Table 4.2: VI=40.4L through VIIIIIII=283L)

CREATE TABLE dravidian_numeral (
    source_code TEXT REFERENCES source(code),
    number INTEGER, compounding_base TEXT, variant TEXT,
    noun_human TEXT, noun_neuter TEXT,
    PRIMARY KEY (source_code, number)) STRICT;
    -- 11 rows (proto-Dravidian numbers 1-10, 100;
    --          McAlpin 1981 via Wells 2015 Table 6.1)
\end{verbatim}
}

\subsection{Codebook schema}

The metrological codebook (\texttt{indus\_codebook.db}) maps Parpola
sign numbers to semantic roles. It is seeded by
\texttt{indus\_seed\_codebook.fsx} and contains seven tables.

{\small
\begin{verbatim}
sign_role       28 rows  sign_id (Parpola) -> role + ref_code
commodity        8 rows  code -> label (JAR GOODS, IRON, etc.)
route            5 rows  code -> label + destination
weight_tier     11 rows  multiplier -> grams + series
quantity_code    6 rows  sign_id -> multiplier
merchant_mark    5 rows  code -> label (UNICORN GUILD, etc.)
positional_rule  3 rows  position -> default_role
\end{verbatim}
}

\noindent
Sign roles divide into five categories: \texttt{commodity} (8~signs),
\texttt{terminal} (5~signs), \texttt{weight} (7~signs),
\texttt{quantity} (6~signs), and \texttt{structural} (2~signs).
The codebook is embedded in the interactive visualisation as
JavaScript \texttt{Map} constants to avoid requiring a second file
upload.

\subsection{Alloy validation report}

Sixteen assertions checked at scope~6. All UNSAT (zero
counterexamples). One valid instance found (SAT).

{\small
\begin{verbatim}
00. check artefactHasOneSite          UNSAT
01. check occurrenceHasOneArtefact    UNSAT
02. check formHasOneParent            UNSAT
03. check noFormOfForm                UNSAT
04. check parallelsCrossSources       UNSAT
05. check noPositionCollision         UNSAT
06. check occurrenceSignExists        UNSAT
07. check formInheritsSource          UNSAT
08. check noSelfParallel              UNSAT
09. check concordanceCrossesSources   UNSAT
10. check concordanceNoDuplicates     UNSAT
11. check cisiSignsValid              UNSAT
12. check measurementOneSample        UNSAT
13. check sampleOneSite               UNSAT
14. check noDupMeasurement            UNSAT
15. check treebankNoDupPosition       UNSAT
16. run   showExample             1/1 SAT
\end{verbatim}
}

\noindent
The Alloy specification (\texttt{spec/indus\_corpus.als}) is
included in the Zenodo artifact. No SQL was written until all
sixteen assertions returned UNSAT.

\subsubsection*{Alloy specs proposed for the new tables}

The 13~tables added in the 2026-05-02 / 2026-05-03 ingests are not
yet covered by the original 16-assertion specification. Twelve
additional assertions are proposed; specs to be authored in a
subsequent revision and run to UNSAT at scope~6 before the
Zenodo update is published.

{\small
\begin{verbatim}
16. check artefactTypeNoOrphanParent      (artefact_type)
17. check artefactTypeAcyclicHierarchy    (artefact_type)
18. check corpusStatOnePerSource          (corpus_stat)
19. check readingDirTotalIsRowSum         (reading_direction_stat)
20. check siteAnchorOnePerSource          (site_book_anchor)
21. check cisiPageNumInRange              (cisi_page)
22. check cisiArtefactIdFormat            (cisi_artefact)
23. check signRoleFunctionExists          (sign_role -> sign_function)
24. check icitInscriptionTextCodeGrammar  (icit_inscription)
25. check signPairFreqNonNegative         (sign_pair_frequency)
26. check signClusterFormat               (sign_cluster)
27. check bibliographyYearInRange         (bibliography_reference)
\end{verbatim}
}

\noindent
\texttt{cisiArtefactIdFormat} requires \texttt{artefact\_id} to
match \texttt{[A-Z]-[0-9]+}. \texttt{icitInscriptionTextCodeGrammar}
requires \texttt{text\_code} to match
$\backslash{\rm +([0-9]\{3\}-)*[0-9]\{3\}\backslash{\rm +}}$
(the canonical ICIT text-code grammar from Fuls 2019 page~7).
\texttt{readingDirTotalIsRowSum} requires the row whose
\texttt{direction} value is \texttt{'TOTAL'} to equal the sum of
the other rows on every numeric column for the same source.

%%% ===================================================================
\section{Reproducibility}\label{sec:repro}

All data and code are deposited under CC~BY~4.0:
\texttt{indus\_corpus.db} (corpus),
\texttt{indus\_codebook.db} (codebook),
\texttt{indus\_lssc.db} (LSSC analysis),
seven F\# scripts (\texttt{.fsx}; see \cref{sec:queries} for the
complete list and usage),
Alloy specification, CMakeLists.txt,
the interactive dashboard (\texttt{ledger-of-meluhha.html}),
and this book.
Any reader with \texttt{sqlite3} can query every claim.
Any reader with \texttt{dotnet fsi} can reproduce every
computation.
The interactive dashboard requires only a modern browser ---
drop \texttt{indus\_corpus.db} onto the page to render the
full trade network, decode two seals live, and filter routes
by commodity.

\subsection{Data update policy}

The codebook and corpus are separated by design.

\paragraph{New corpus data} (additional CISI seals, new
archaeological sites, new morphological parallels) requires only
re-running the ingest pipeline:
\texttt{cmake --build build --target corpus}.
No script, codebook, or book change is needed.

\paragraph{New codebook entries} (a new commodity, route, or sign
role assignment) propagate through a four-step chain:
\begin{enumerate}[nosep]
\item Edit \texttt{indus\_seed\_codebook.fsx} (the single source
  of truth).
\item Re-seed: \texttt{cmake --build build --target codebook}.
\item Re-generate typed records: \texttt{cd hydra \&\& sqlhydra
  sqlite}.
\item Update the frozen F\# types in
  \texttt{IndusCodebookTypes.fsx} and the embedded JavaScript maps
  in \texttt{ledger.ts}.
\end{enumerate}
The codebook has been stable at 28~sign roles, 8~commodities, and
5~routes since the initial seeding. By construction the open
inventory (merchant marks, artefact IDs) lives in the corpus, not
the codebook.

\subsubsection*{Coding standard for ingest scripts}

Every script that writes to \texttt{indus\_corpus.db} is named
\texttt{*\_to\_sqlite.fsx} and lives at the repository root. This
convention encodes the read/write split that the F\# coding
standard (\texttt{spec/fsharp/reference/fsharp\_coding\_standard.tex})
requires: write-path scripts may use raw SQL through
\texttt{Microsoft.Data.Sqlite}; read-path scripts must consume the
database via SqlHydra-generated typed records (no
\texttt{SqliteCommand}, \texttt{ExecuteReader}, or
\texttt{GetString} calls). The \texttt{*\_to\_sqlite.fsx} suffix is
the file-naming half of the same rule.

Ingest scripts share four further conventions:

\begin{itemize}[nosep]
\item \texttt{\#load "lib/shell/Shell.fsx"} for any process
  spawning. \texttt{lib/shell/Shell.fsx} is an absolute symlink to
  \texttt{\textasciitilde{}/fsx-tools/lib/Shell.fsx}, the canonical
  shared F\# DSL for grep/find/sed/zsh-style operations and
  process invocation. \texttt{Shell.runAsync} replaces raw
  \texttt{ProcessStartInfo} usage.
\item Header citation pair: \texttt{// Implements: Ledger of
  Meluhha §<section> --- <what>} and
  \texttt{// Coding standard: spec/fsharp/reference/fsharp\_coding\_standard.tex}
  immediately under the SPDX line.
\item Records over tuples (T1): hardcoded data tables that mirror
  paper tables are typed F\# records, not tuples. The record name
  matches the table label in the paper.
\item Idempotent UPSERT: every \texttt{INSERT} carries
  \texttt{ON CONFLICT \ldots DO UPDATE} so re-running a script
  against an already-populated database is a no-op rather than a
  duplicate-row error.
\end{itemize}

%%% ===================================================================
\clearpage
\section{Interactive Dashboard User Guide}\label{sec:dashboard}

The file \texttt{ledger-of-meluhha.html} is a self-contained
interactive visualisation of the entire trade network, decoded
seals, and codebook. It requires no server, no build chain, and no
framework --- only a modern browser and the corpus database file.

\subsection{Getting started}

\begin{enumerate}
\item Open \texttt{ledger-of-meluhha.html} in any modern browser
  (Chrome, Firefox, Safari, Edge).
\item The landing page shows the title and a dashed drop zone:
  \textit{``drag indus\_corpus.db to begin.''}
\item Drag \texttt{indus\_corpus.db} from any folder onto the drop
  zone. The file is loaded entirely in-browser via sql.js (WASM).
  No data leaves your machine.
\item The dashboard renders immediately.
\end{enumerate}

\subsection{Layout}

The dashboard is a three-column grid:

\paragraph{Left panel --- Decoded seals.}
Two seal cards for M-52A and M-148A. Each card shows the five-field
decode: F1~(merchant mark), F2~(commodity), F3~(weight tier),
F4~(quantity), F5~(route). Parpola sign numbers and concordance
paths are displayed in \texttt{JetBrains Mono}; resolved values in
\textit{Cormorant Garamond}. Below the cards, the Alloy
verification badge: \textit{``16 ASSERTIONS UNSAT AT SCOPE~6.''}
A radiometric dating panel shows calibrated dates from the corpus.

\paragraph{Center panel --- Trade network map.}
A Leaflet map centered on the IVC trade zone, showing 22
archaeological sites across four categories:

\begin{itemize}
\item IVC sites (Mohenjo-daro, Harappa, Lothal, Kalibangan,
  Dholavira, Rojdi, Shortughai)
\item Tamil Nadu supply sites (Kodumanal, Keeladi, Thulukarpatti,
  Karur, Teriruveli, Adichanallur, Sivagalai)
\item Mesopotamian destinations (Ur, Kish, Tell Asmar, Susa)
\item Gulf hubs (Dilmun/Bahrain, Magan/Oman)
\item Resource sites (Khetri copper mines)
\end{itemize}

Trade routes are drawn as curved polylines. Route stroke width
is proportional to relative trade volume derived from cuneiform
Meluhha records. Hover over any route to see the commodity,
sign number, and evidence basis in the tooltip.

\paragraph{Right panel --- Reference charts.}
Three stacked panels:

\begin{enumerate}
\item \textbf{Sign frequency} --- D3 bar chart of cumulative sign
  distribution. A dashed TBA-red line marks the 80\% threshold at
  67~signs. Hover over any bar for the exact count.
\item \textbf{Weight codebook} --- horizontal bars for all 11 weight
  tiers. Binary tiers (clay colour) and decimal tiers (TBA-red)
  are visually distinguished. Each bar shows the gram value.
\item \textbf{Commodity manifest} --- clickable table of 8
  commodity classes with sign number, name, primary route, and
  corpus share. Click a row to highlight the corresponding trade
  route on the map; click again to deselect. The active row is
  marked in TBA-red.
\end{enumerate}

\subsection{Themes}

Four accessibility themes are available via the coloured dots in the
upper-right corner:

\begin{longtable}{@{}llp{7cm}@{}}
\toprule
Theme & Background & Notes \\
\midrule
\endfirsthead
\bottomrule
\endlastfoot
Clay (default) & Dark (\texttt{\#0f0c07}) & Warm sand text, clay data, grain texture. Designed for extended reading in low ambient light. \\
\midrule
Parchment & Light (\texttt{\#f5f0e4}) & High-contrast dark text on cream. For projection, printing, or bright-room reading. \\
\midrule
High Contrast & Black (\texttt{\#000000}) & WCAG~AAA compliant. Yellow data on black. No grain texture. For visually impaired users. \\
\midrule
Seal & Dark blue-grey (\texttt{\#12161e}) & Reduced eye strain with blue-grey tones. For long sessions. \\
\end{longtable}

\noindent
Theme selection persists for the session. All data visualisations
adapt to the selected theme.

\subsection{How the decode works live}

When the database loads, the dashboard executes the full decode
chain in real time:

\begin{enumerate}[nosep]
\item Query \texttt{cisi\_inscription} for M-52A and M-148A sign
  sequences.
\item For each Parpola sign ID, query \texttt{sign\_concordance} to
  resolve the Mahadevan number.
\item Look up the Mahadevan number in the embedded
  \texttt{signRoleMap} (the codebook, frozen as JavaScript
  \texttt{Map} constants).
\item Resolve the role to a commodity name, weight tier, or route
  destination via \texttt{commodityMap} and \texttt{routeMap}.
\item Render the five-field card.
\end{enumerate}

\noindent
The decode is deterministic and reproducible. The same database
produces the same cards every time. No server-side computation.
No randomness. No model.

\subsection{What the user can do}

\begin{itemize}
\item \textbf{Pan and zoom} the map to explore site locations and
  route geometries.
\item \textbf{Hover} over trade routes to see commodity, evidence,
  and sign numbers.
\item \textbf{Click} commodity rows to filter the map to a single
  trade corridor.
\item \textbf{Switch themes} for accessibility or presentation.
\item \textbf{Inspect the decode} --- every sign number, concordance
  path, and codebook lookup is visible on the seal cards.
  Nothing is hidden.
\end{itemize}

\subsection{Technical requirements}

\begin{itemize}
\item Any modern browser with WebAssembly support (all major
  browsers since 2017).
\item The \texttt{indus\_corpus.db} file (produced by
  \texttt{indus\_ingest\_corpus.fsx} and
  \texttt{indus\_ingest\_cisi.fsx}).
\item Internet connection for CDN resources (sql.js WASM, D3.js,
  Leaflet, Google Fonts). For fully offline use, bundle the
  dependencies locally.
\item No installation. No build step. No server.
\end{itemize}

%%% ===================================================================
\clearpage
\section{Reproducing Every Claim}\label{sec:queries}

Every claim in this book is reproduced by an F\# script. The
scripts are the source of truth. They read from SQLite, perform
the computation, and print the result. No raw SQL is required ---
the scripts embed it.

\subsection{The thirteen scripts}

The original seven scripts produce the analytical claims of the book.
Six write-path \texttt{*\_to\_sqlite.fsx} scripts shipped 2026-05-02 /
2026-05-03 ingest the secondary sources discussed in §The Field /
Prior art into \texttt{indus\_corpus.db}.

\begin{longtable}{@{}lp{8cm}@{}}
\toprule
Script & What it reproduces \\
\midrule
\endfirsthead
\bottomrule
\endlastfoot
\multicolumn{2}{@{}l@{}}{\textit{Original analytical scripts (7).}} \\
\midrule
\texttt{indus\_ingest\_corpus.fsx}
  & Builds \texttt{indus\_corpus.db} from four sources
    (Mahadevan, Rajan--Sivanantham, Tamil Treebank,
    radiometric dating). \\
\midrule
\texttt{indus\_ingest\_cisi.fsx}
  & Ingests 396~Parpola sign definitions, 179~CISI inscriptions,
    1{,}003~sign occurrences, and 334~Mahadevan concordance
    mappings from the mayig JSON corpus into
    \texttt{indus\_corpus.db}. Spot-checks M-52A decode. \\
\midrule
\texttt{indus\_seed\_codebook.fsx}
  & Builds \texttt{indus\_codebook.db}: 28~sign roles,
    8~commodities, 5~routes, 11~weight tiers, 6~quantity codes,
    5~merchant marks, 3~positional rules. \\
\midrule
\texttt{indus\_decoder.fsx}
  & Full decode chain for M-52A and M-148A: CISI inscription
    $\to$ concordance $\to$ codebook $\to$ field assignment.
    Prints the five-field cargo tag. \\
\midrule
\texttt{indus\_tn\_decoder.fsx}
  & Tamil Nadu cross-corpus decode: 20~morphological parallels
    through the concordance bridge. \\
\midrule
\texttt{indus\_bernoulli.fsx}
  & Five-path Bernoulli independence test ($p = 3.125 \times
    10^{-5}$), Monte Carlo validation ($10^7$ trials),
    Fisher's combined test, Bayesian update, sensitivity
    analysis. \\
\midrule
\texttt{indus\_lssc.fsx}
  & Transition entropy and LSSC (Latent Structural State
    Contraction) analysis on the Tamil proxy corpus.
    Closure/option sign classification. \\
\midrule
\texttt{indus\_bigram.fsx}
  & Mass decode of all 179~inscriptions (117~commodity hits,
    37~route hits, South terminal = 0). Bigram structure
    analysis: mapped vs unmapped signs. Conclusion: no
    alphabet detected (1.1~successors/sign, 71\% unique
    bigrams = open inventory, not phonetic layer). \\
\midrule
\multicolumn{2}{@{}l@{}}{\textit{Secondary-source ingest scripts (6, 2026-05-02 / 03).}} \\
\midrule
\texttt{indus\_fuls2022\_to\_sqlite.fsx}
  & Fuls 2022 \textit{Corpus of Indus Inscriptions} metadata
    scaffold: source row, 73~sites with book-page anchors,
    artefact typology (10 top-level + 23 subtypes),
    corpus headline statistics (4660 / 5644 / 19831), reading
    direction stats (Table~2.5). Phase~1 ingest of the published
    35-page methodology PDF; the gated 570-page Chapter~3 site
    sections are not extractable. \\
\midrule
\texttt{cisi\_ocr\_to\_sqlite.fsx}
  & Joshi--Parpola 1987 Vol.~1: 862-page scanned PDF rendered at
    300dpi via pdftoppm, OCR'd via tesseract \texttt{--psm 6},
    parsed via regex into \texttt{cisi\_page} (862 rows) and
    \texttt{cisi\_artefact} (3305 rows, 1399 unique IDs across
    six site prefixes). 91\% recall against the documented
    1{,}537 artefacts. \\
\midrule
\texttt{cisi\_ocr\_cleanup\_to\_sqlite.fsx}
  & Post-process for the OCR pipeline: site-name typo
    normalisation (MOHENIJO-DARO, HARAPPAS, HULASI, GRAFFITI/SEALS
    prefixes); \texttt{is\_phantom} marker on
    \texttt{cisi\_artefact} rows whose page lacks a canonical
    Indus-site header. Removes phantom rows from the
    introduction-text and index-page noise. \\
\midrule
\texttt{icit\_help\_to\_sqlite.fsx}
  & Fuls 2019 ICIT documentation PDF: 16~sign-function codes,
    4~reading-direction codes, 10~real Wells-numbered ICIT
    inscriptions (the only such data outside the unreachable
    \texttt{caddy.igg.tu-berlin.de}), 40~sign-role assignments,
    20~sign-pair frequencies (12 TMK--TMK + 8 sign~700~$\times$~NUM),
    4~paradigmatic sign clusters, 8~bibliography references. \\
\midrule
\texttt{wells2015\_to\_sqlite.fsx}
  & Wells 2015 \textit{Archaeology and Epigraphy of Indus
    Writing} Phase~1 (prose-extractable): 5~new sign-function
    codes (INS, ICTM, EMS, MMS, LMS), 46~sign-role assignments
    including the canonical ICTM identification (signs 1, 2, 60),
    4~Harappa volumetric units, 6~bibliography entries
    (Wells 2006, Yadav 2008, Knorozov 1968, Marshall 1931,
    Mackay 1938, Hemmy 1931). \\
\midrule
\texttt{wells2015\_tables\_to\_sqlite.fsx}
  & Wells 2015 Phase~2 (table OCR): 3~more volumetric units
    (full Harappa series VI=40.4L through VIIIIIII=283.0L now
    7~rows), Wells's own sign~700~$\times$~Long-Linear-Stroke
    matrix from Table~5.3 (3~rows; differ slightly from Fuls
    2019 figures, both kept), 11~more sign-role assignments,
    11~proto-Dravidian numerals from Table~6.1 (McAlpin 1981
    reproduced). \\
\end{longtable}

\subsection{Running a script}

Every script follows the same pattern:

\begin{verbatim}
dotnet fsi <script> <input-args>
\end{verbatim}

\noindent
Examples:

\begin{verbatim}
# Build the corpus from source data
dotnet fsi indus_ingest_corpus.fsx \
  ~/indus-valley indus_corpus.db

# Ingest CISI from the mayig fork
dotnet fsi indus_ingest_cisi.fsx \
  ~/indus-valley-script-corpus indus_corpus.db

# Mass decode + bigram analysis
dotnet fsi indus_bigram.fsx \
  indus_corpus.db indus_codebook.db

# Bernoulli independence test
dotnet fsi indus_bernoulli.fsx
\end{verbatim}

\noindent
Each script prints its results to stdout. No intermediate files
are required. The only dependency is \texttt{dotnet fsi} (the
F\# interactive compiler, included with the .NET SDK) and
\texttt{Microsoft.Data.Sqlite} (pulled automatically via NuGet
on first run).

\subsection{Why F\# scripts, not raw SQL}

SQL is embedded inside the F\# scripts. The scripts are the
reproducibility layer because:

\begin{itemize}
\item A script is a \emph{complete computation}: it opens the
  database, runs the query, processes the results, and prints
  the conclusion. A raw SQL query produces a table that still
  requires interpretation.
\item Scripts are versioned in git. Every commit that changes a
  claim also changes the script that produces it.
\item Scripts can be composed: \texttt{indus\_bigram.fsx} runs the
  mass decode, the commodity distribution, the route distribution,
  and the bigram analysis in a single invocation.
\item Scripts carry the interpretation: the ``CONCLUSION: No
  alphabet detected'' line is printed by the script, not inferred
  by the reader from a table of numbers.
\end{itemize}

\subsection{Excel / Google Sheets workflow}

For researchers who prefer spreadsheets, run the CSV export
script:

\begin{verbatim}
# Export via the build system
cmake --build build --target csv-export
\end{verbatim}

\noindent
This produces three CSV files:
\texttt{cisi\_inscriptions.csv},
\texttt{concordance.csv}, and
\texttt{codebook\_roles.csv}.

To reproduce the decode chain in Excel:

\begin{enumerate}
\item Open \texttt{cisi\_inscriptions.csv}. Find M-52A. The
  \texttt{signs} column contains:
  \texttt{P324,P172,P145,P050,...}
\item Open \texttt{concordance.csv}. Filter
  \texttt{parpola\_id} to P324.
  Result: \texttt{mahadevan\_ids}~= M342.
\item Open \texttt{codebook\_roles.csv}. Filter
  \texttt{sign\_id}~= 342.
  Result: \texttt{role}~= commodity,
  \texttt{ref\_code}~= jar.
\item Decode complete: P324~$\to$~M342~$\to$~jar goods.
\item Use \texttt{VLOOKUP} or \texttt{INDEX/MATCH} across the
  three sheets to automate the chain for all 179~inscriptions.
\end{enumerate}

\subsection{Other languages}

The databases are standard SQLite files. Any language with a
SQLite driver can query them:

\begin{itemize}
\item \textbf{Python:} \texttt{import sqlite3} (standard library).
\item \textbf{R:} \texttt{library(RSQLite)}.
\item \textbf{JavaScript:} sql.js (WASM, as used in the dashboard).
\end{itemize}

\noindent
But the F\# scripts are the canonical form. If the script says
``117 commodity hits'' and your SQL says 116, the script is right.

%%% ===================================================================
\clearpage
\section{Plain-Language Summary}

\textit{For journalists, educators, and general readers.}

\subsection{What is this about?}

The Indus Valley Civilisation (c.\ 2600--1900~\textsc{bce}) was
one of the largest Bronze Age societies --- bigger than Egypt and
Mesopotamia combined. They left about 4{,}000 objects with short
inscriptions. For a hundred years, people have tried to read these
as a lost language. Nobody has succeeded.

Imagine archaeologists in the year 6000~\textsc{ce} finding a
shipping container with a barcode sticker and no scanner. They
might spend centuries trying to read the barcode as English. It is
not English. It is a structured record. The Indus seals are the
Bronze Age version of that sticker.

\subsection{What does this book argue?}

The inscriptions are shipping labels, not language. Each seal
records:

\begin{enumerate}
\item \textbf{Who} is shipping (animal picture = guild logo).
  \textit{The unicorn motif on hundreds of seals with different
  sign sequences = same company, different shipments.}
\item \textbf{What} commodity (the ``jar'' sign on 10\% of all
  inscriptions = dominant shipping container, like seeing
  ``container'' on most bills of lading).
\item \textbf{How much} it weighs (the seal points at the weight
  object in the set. Weights accurate to 0.5\% across the entire
  civilisation).
\item \textbf{How many} units.
\item \textbf{Where} it is going (Mesopotamia, Bahrain, Oman, or
  internal routes).
\end{enumerate}

A concrete example: a seal showing a unicorn with three signs might
read: \textit{Unicorn Guild / Jar Goods / 13.7g tier /
Mesopotamia.}

\subsection{What did we find?}

We decoded two real seals from a published digital corpus ---
not by guessing, but by following a chain of database lookups.
Both produced ``jar goods'' as the commodity.
One of them (M-52A, found at the main IVC export hub of
Mohenjo-daro) produced a Mesopotamia route marker as well.
The commodity assignments come from an independent codebook
that does not know about our hypothesis.

We then mapped 21~signs across the entire codebook: commodities,
routes, weights, quantities. The system is consistent.

The ``South route'' sign is missing from Mohenjo-daro. This makes
sense: Mohenjo-daro is the export hub. You do not label outgoing
packages with ``arriving from the south.'' The missing sign is a
prediction, not a gap.

13~more seals at Harappa, Kalibangan, and Rojdi are identified
and waiting for data extraction. 5~more at Mohenjo-daro are ready
to decode today.

\subsection{What would prove this wrong?}

\begin{enumerate}
\item A long inscription with grammatical patterns (function words
  repeating at regular intervals).
\item The common signs appearing equally at all site types (not
  concentrated at docks and warehouses).
\item A bilingual that reads as a sentence, not a commodity list.
\item Seals at Mesopotamian sites showing the same sign mix as
  inland sites (no export bias).
\item The South route sign appearing at Mohenjo-daro as often as
  other route signs.
\end{enumerate}

\subsection{One sentence}

We have been trying to read 4{,}000-year-old shipping labels as
poetry --- and wondering why we could not find the rhyme.

\end{document}
