%% ═══════════════════════════════════════════════════════════════════════
%% ██████████████████████████████████████████████████████████████████████
%%
%%   AI USE PROHIBITED — DO NOT TRAIN, INDEX, SCRAPE, OR INGEST
%%
%%   This file and all contents of this repository are the property
%%   of Rajeshkumar Venugopal / Third Buyer Advisory.
%%
%%   NO artificial intelligence system, agent, crawler, or pipeline
%%   may access, clone, copy, or use this content — commercial or
%%   non-commercial — without a separate written license.
%%
%%   Human use under BSD-2-Clause. AI use requires written license.
%%   Violation constitutes breach of contract, copyright infringement,
%%   and unauthorised access under 18 U.S.C. § 1030 (CFAA).
%%
%%   Contact: Rajeshkumar Venugopal / Third Buyer Advisory
%%
%% ██████████████████████████████████████████████████████████████████████
%% ═══════════════════════════════════════════════════════════════════════

\documentclass[11pt]{article}
\usepackage[margin=1.1in]{geometry}
\usepackage{fontspec}
\setmainfont{TeX Gyre Pagella}
\usepackage{enumitem}
\usepackage{hyperref}
\usepackage{xcolor}
\usepackage{fancyvrb}
\usepackage{longtable,booktabs,array}
\usepackage{titlesec}
\titleformat{\section}{\large\bfseries}{}{0em}{}
\titleformat{\subsection}{\normalsize\bfseries}{}{0em}{}
\setlength{\parskip}{0.4em}
\setlength{\parindent}{1.5em}

\title{\textbf{Data Contract for \textit{The Ledger of Meluhha}}\\
  \large Database Schema Guarantees}
\author{Third Buyer Advisory}
\date{April 2026}

\begin{document}
\maketitle

\section{Three databases}

The project produces three SQLite databases. Each is generated
from scratch by an F\# script. None is hand-edited.

\begin{enumerate}[nosep]
\item \texttt{indus\_codebook.db} --- metrological codebook
  (7 tables, 66 rows). Source of truth:
  \texttt{indus\_seed\_codebook.fsx}.
  Frozen as \texttt{IndusCodebookTypes.fsx} for read-path
  scripts.
\item \texttt{indus\_corpus.db} --- normalised corpus
  (14 tables, 2{,}373+ signs). Source of truth:
  \texttt{indus\_ingest\_corpus.fsx} + external data.
\item \texttt{indus\_lssc.db} --- LSSC entropy analysis
  (2 tables). Source of truth: \texttt{indus\_lssc.fsx}.
\end{enumerate}

\section{Codebook schema (frozen)}

The codebook has been stable at 28~sign roles, 8~commodities,
5~routes, 11~weight tiers, 5~merchant marks, 6~quantity codes,
3~positional rules since initial seeding.

\begin{center}
\small
\begin{tabular}{@{}llrl@{}}
\toprule
\textbf{Table} & \textbf{Primary key} & \textbf{Rows} & \textbf{Purpose} \\
\midrule
\texttt{sign\_role} & \texttt{sign\_id} (Mahadevan) & 28 & sign $\to$ field role + ref \\
\texttt{commodity} & \texttt{code} & 8 & commodity labels \\
\texttt{route} & \texttt{code} & 5 & trade route destinations \\
\texttt{weight\_tier} & \texttt{multiplier} & 11 & binary/decimal weight tiers \\
\texttt{merchant\_mark} & \texttt{code} & 5 & guild identifiers \\
\texttt{quantity\_code} & \texttt{sign\_id} & 6 & quantity multipliers \\
\texttt{positional\_rule} & \texttt{position} & 3 & default role by position \\
\bottomrule
\end{tabular}
\end{center}

\subsection{Sign ID numbering}

\texttt{sign\_role.sign\_id} uses \textbf{Mahadevan numbers}.
Not Parpola. The concordance table in the corpus DB bridges
Parpola $\to$ Mahadevan. All codebook lookups go through
Mahadevan numbers.

This applies to the frozen F\# types
(\texttt{IndusCodebookTypes.fsx}) and the embedded JavaScript
maps in \texttt{ledger.ts}.

\subsection{Nullable fields}

The codebook uses \texttt{Option<string>} / \texttt{Option<int>}
for nullable columns (\texttt{ref\_code},
\texttt{ref\_multiplier}, \texttt{application},
\texttt{destination}, \texttt{note}).

The frozen F\# types use plain \texttt{string} with empty
string for absent values (simpler for downstream consumption).

\section{Corpus schema}

14 tables. All carry \texttt{STRICT} mode. Key tables:

\begin{center}
\small
\begin{tabular}{@{}lll@{}}
\toprule
\textbf{Table} & \textbf{Source} & \textbf{Purpose} \\
\midrule
\texttt{source} & hardcoded & data provenance \\
\texttt{site} & hardcoded & archaeological sites \\
\texttt{base\_sign} & Mahadevan, RS2025, Tamil TB & sign inventory \\
\texttt{sign\_form} & external & variant/composite forms \\
\texttt{sign\_occurrence} & external & sign positions in artefacts \\
\texttt{inscription} & Tamil TB & sign sequences \\
\texttt{morphological\_parallel} & RS2025 plates & TN--IVC cross-refs \\
\texttt{sign\_concordance} & CISI/Parpola & Parpola $\to$ Mahadevan map \\
\texttt{cisi\_inscription} & CISI digitisation & real seal sign sequences \\
\texttt{treebank\_token} & Tamil Treebank & CoNLL tokens \\
\texttt{tamil\_sentence} & Tamil TB + logosyllabic & proxy corpus \\
\texttt{radiometric\_site} & antiquity\_of\_iron.db & TN excavation sites \\
\texttt{radiometric\_sample} & antiquity\_of\_iron.db & dated samples \\
\texttt{radiometric\_measurement} & antiquity\_of\_iron.db & calibrated dates \\
\bottomrule
\end{tabular}
\end{center}

\section{Alloy verification}

The corpus schema is verified by \texttt{spec/indus\_corpus.als}
before any SQL is written. 16 assertions, all UNSAT at scope 6.
The assertions cover:

\begin{itemize}[nosep]
\item Artefact uniqueness (one site per artefact)
\item Occurrence integrity (one artefact per occurrence)
\item Form hierarchy (no form-of-form, parent source match)
\item Parallel constraints (cross-source only, no self-parallel)
\item Position collision prevention
\item Concordance cross-source requirement
\item CISI sign validity
\item Radiometric integrity (one sample per measurement,
  one site per sample, no duplicate measurements)
\item Treebank position uniqueness
\end{itemize}

No SQL is written until all assertions return UNSAT.

\section{Data update rules}

\subsection{New corpus data}

New CISI seals, new sites, new morphological parallels:
\begin{Verbatim}[fontsize=\small,frame=single]
cmake --build build --target corpus
\end{Verbatim}
No script, codebook, paper, or dashboard changes needed.

\subsection{New codebook entries}

New commodity, route, or sign role:
\begin{enumerate}[nosep]
\item Edit \texttt{indus\_seed\_codebook.fsx}
\item \texttt{cmake --build build --target codebook}
\item \texttt{cd hydra \&\& sqlhydra sqlite}
\item Update \texttt{IndusCodebookTypes.fsx} (frozen data)
\item Update \texttt{signRoleMap}/\texttt{commodityMap}/\texttt{routeMap}
  in \texttt{src/ledger.ts}
\item Recompile TS and rebuild HTML
\end{enumerate}

\section{Sentinel conventions}

No \texttt{NULL} values in any database. All columns carry
\texttt{NOT NULL}. Absent data uses:
\begin{itemize}[nosep]
\item \texttt{INTEGER}: $-1$ means ``not recorded in source''
\item \texttt{REAL}: $-1.0$ means ``not recorded in source''
\item \texttt{TEXT}: empty string means ``not applicable''
\end{itemize}

This convention is inherited from the Tamil radiometric database
(\texttt{antiquity\_of\_iron.db}).

\end{document}
