%% ═══════════════════════════════════════════════════════════════════════
%% ██████████████████████████████████████████████████████████████████████
%%
%%   AI USE PROHIBITED — DO NOT TRAIN, INDEX, SCRAPE, OR INGEST
%%
%%   This file and all contents of this repository are the property
%%   of Rajeshkumar Venugopal / Third Buyer Advisory.
%%
%%   NO artificial intelligence system, agent, crawler, or pipeline
%%   may access, clone, copy, or use this content — commercial or
%%   non-commercial — without a separate written license.
%%
%%   Human use under BSD-2-Clause. AI use requires written license.
%%   Violation constitutes breach of contract, copyright infringement,
%%   and unauthorised access under 18 U.S.C. § 1030 (CFAA).
%%
%%   Contact: Rajeshkumar Venugopal / Third Buyer Advisory
%%
%% ██████████████████████████████████████████████████████████████████████
%% ═══════════════════════════════════════════════════════════════════════

\documentclass[11pt]{article}
\usepackage[margin=1.1in]{geometry}
\usepackage{fontspec}
\setmainfont{TeX Gyre Pagella}
\usepackage{enumitem}
\usepackage{hyperref}
\usepackage{xcolor}
\usepackage{fancyvrb}
\usepackage{titlesec}
\titleformat{\section}{\large\bfseries}{}{0em}{}
\titleformat{\subsection}{\normalsize\bfseries}{}{0em}{}
\setlength{\parskip}{0.4em}
\setlength{\parindent}{1.5em}

\title{\textbf{TypeScript Coding Standard for \textit{The Ledger of Meluhha}}\\
  \large Dashboard and Interactive Visualization}
\author{Third Buyer Advisory}
\date{April 2026}

\begin{document}
\maketitle

\section{Delivery model}

The dashboard ships as a \textbf{single HTML file}
(\texttt{ledger-of-meluhha.html}) with all JavaScript inlined.
No bundler, no node\_modules, no build server.
The user drops \texttt{indus\_corpus.db} onto the page.

Build chain: \texttt{src/ledger.ts} $\to$ \texttt{npx tsc}
$\to$ \texttt{dist/ledger.js} $\to$ paste into HTML
\texttt{<script>} block.

\section{External dependencies (CDN only)}

Four CDN-loaded libraries. No local copies.

\begin{enumerate}[nosep]
\item \texttt{sql.js} --- SQLite WASM (loads \texttt{.db} in browser)
\item \texttt{Leaflet} --- map tiles + polylines
\item \texttt{D3.js} --- frequency chart
\item Google Fonts --- Cormorant Garamond, JetBrains Mono
\end{enumerate}

Map tiles require internet (OpenTopoMap). No offline mode.

\section{Codebook embedding}

The codebook tables (\texttt{sign\_role}, \texttt{commodity},
\texttt{route}) are embedded as typed JavaScript \texttt{Map}
constants. These \textbf{must match}
\texttt{IndusCodebookTypes.fsx} exactly --- same sign IDs,
same codes, same values. The sign IDs are \textbf{Mahadevan
numbers}, not Parpola numbers.

When the frozen F\# types change, update:
\begin{itemize}[nosep]
\item \texttt{signRoleMap} (28 entries)
\item \texttt{commodityMap} (8 entries)
\item \texttt{routeMap} (5 entries)
\item Weight array (11 tiers)
\end{itemize}

\section{Seal decode in the browser}

The dashboard decodes seals by parsing Parpola sign IDs from
\texttt{cisi\_inscription} rows in the dropped corpus DB,
then looking up the Mahadevan number in the embedded
\texttt{signRoleMap}. The decode chain must match the paper's
description in \S6.2.

The seal IDs shown in the dashboard must match the seals
discussed in the paper. Currently: M-52A and M-148A.

\section{CSS rules}

\begin{itemize}[nosep]
\item Four themes: clay (dark default), parchment (light),
  high contrast (WCAG AAA), seal (blue-grey).
\item All colours via CSS custom properties (\texttt{--clay},
  \texttt{--sand}, \texttt{--tba-red}, etc.).
\item Do not apply CSS \texttt{width}/\texttt{height} to
  Leaflet's internal SVG elements. Leaflet manages its own
  SVG overlay dimensions.
\item Commodity table rows have \texttt{cursor:pointer} and
  \textbf{must} have click handlers that highlight matching
  trade routes on the map.
\end{itemize}

\section{Route rendering}

Trade routes render as Leaflet polylines with quadratic Bezier
curves. Four cases handled:
\begin{enumerate}[nosep]
\item \texttt{from} $\to$ \texttt{to} (direct)
\item \texttt{from} $\to$ \texttt{via} $\to$ \texttt{to} (multi-segment)
\item \texttt{from} $\to$ \texttt{via} (terminates at via)
\item \texttt{from} only (skip --- no geometry)
\end{enumerate}

Routes render before sites (z-order: routes under, sites on top).
Polyline refs stored in \texttt{routeLines[]} for commodity filtering.

\section{Weight labels}

Weight application labels in the JS must match the seed data
in \texttt{indus\_seed\_codebook.fsx} verbatim. No abbreviations.

\end{document}
