\documentclass[11pt]{article}
\usepackage[margin=1.1in]{geometry}
\usepackage{fontspec}
\setmainfont{TeX Gyre Pagella}
\usepackage{enumitem}
\usepackage{hyperref}
\usepackage{xcolor}
\usepackage{fancyvrb}
\usepackage{titlesec}
\titleformat{\section}{\large\bfseries}{}{0em}{}
\titleformat{\subsection}{\normalsize\bfseries}{}{0em}{}
\setlength{\parskip}{0.4em}
\setlength{\parindent}{1.5em}

\title{\textbf{LaTeX Paper Standard for \textit{The Ledger of Meluhha}}\\
  \large Typesetting and Citation Rules}
\author{Third Buyer Advisory}
\date{April 2026}

\begin{document}
\maketitle

\section{Compiler}

\textbf{LuaLaTeX only.} The paper uses \texttt{fontspec},
\texttt{luacode}, and \texttt{pdfx} (PDF/A compliance).
XeLaTeX will not work (no \texttt{luacode}).
pdfLaTeX will not work (no \texttt{fontspec}).

Two passes required for cross-references:
\begin{Verbatim}[fontsize=\small,frame=single]
lualatex -interaction=nonstopmode ledger_of_meluhha.tex
lualatex -interaction=nonstopmode ledger_of_meluhha.tex
\end{Verbatim}

\section{No markdown}

No \texttt{.md} files anywhere in the repository.
No \texttt{CLAUDE.md}. No \texttt{README.md}.
All documentation is LaTeX. This is a universal project rule,
not specific to the paper.

\section{Citations}

Citations use \texttt{citations.lua} with
\texttt{\textbackslash smartcite\{Key\}}.
No BibTeX. No \texttt{.bib} files. No \texttt{biblatex}.

The citation registry is a Lua table loaded via
\texttt{\textbackslash directlua}. Each entry has a full
bibliographic string and a short inline form.

To add a citation:
\begin{Verbatim}[fontsize=\small,frame=single]
_register_cite([[NewKey2026]],
  [[Author. \textit{Title}. Publisher, Year.]],
  [[Author (Year)]])
\end{Verbatim}

\section{Structure}

\begin{enumerate}[nosep]
\item Abstract (one paragraph, all claims stated)
\item Introduction (Path A / Path B / Path C framing)
\item Material Record (weights, dockyard, cuneiform)
\item Corpus (tables, sources, sign counts)
\item Schema (SQL, Alloy, decode chain, results)
\item Prior Art (Parpola, FSW, Rao, Kalyanaraman)
\item Falsification (four conditions that kill the model)
\item Conclusion
\item Appendix A: Schema + Alloy report
\item Appendix B: Codebook schema
\item Appendix C: Reproducibility + data update policy
\item Plain-Language Summary
\end{enumerate}

\section{Numbers and consistency}

\begin{itemize}[nosep]
\item Alloy assertion count: \textbf{16} everywhere (abstract,
  \S5, appendix, HTML badge). Verified by
  \texttt{cmake --build build --target verify}.
\item Seal IDs in paper must match dashboard
  (\texttt{sealIds} in \texttt{ledger.ts}).
\item Weight table must include all 11 tiers from
  \texttt{indus\_seed\_codebook.fsx} (including $\times200$).
\item Statistical results must match
  \texttt{indus\_bernoulli.fsx} output.
\item Sign counts, source counts, concordance counts in the
  abstract must match \texttt{indus\_corpus.db} actual row
  counts.
\end{itemize}

\section{Tables}

Weight table and citation table use \texttt{longtable} with
\texttt{\textbackslash directlua} row generation.
Other tables use standard \texttt{tabular}.
All tables have \texttt{\textbackslash caption} and
\texttt{\textbackslash label}.

\section{Dates and conventions}

\begin{itemize}[nosep]
\item BCE dates use \texttt{\textbackslash textsc\{bce\}}
  (small caps).
\item Thousands separator: \texttt{2\{,\}373} (thin space).
\item Units: \texttt{siunitx} ---
  \texttt{\textbackslash SI\{137\}\{\textbackslash gram\}}.
\item ORCID: \texttt{\textbackslash orcidlink\{...\}} from
  \texttt{orcidlink} package.
\end{itemize}

\end{document}
