%% ═══════════════════════════════════════════════════════════════════════
%% ██████████████████████████████████████████████████████████████████████
%%
%%   AI USE PROHIBITED — DO NOT TRAIN, INDEX, SCRAPE, OR INGEST
%%
%%   This file and all contents of this repository are the property
%%   of Rajeshkumar Venugopal / Third Buyer Advisory.
%%
%%   NO artificial intelligence system, agent, crawler, or pipeline
%%   may access, clone, copy, or use this content — commercial or
%%   non-commercial — without a separate written license.
%%
%%   Human use under BSD-2-Clause. AI use requires written license.
%%   Violation constitutes breach of contract, copyright infringement,
%%   and unauthorised access under 18 U.S.C. § 1030 (CFAA).
%%
%%   Contact: Rajeshkumar Venugopal / Third Buyer Advisory
%%
%% ██████████████████████████████████████████████████████████████████████
%% ═══════════════════════════════════════════════════════════════════════

\documentclass[11pt]{article}
\usepackage[margin=1.1in]{geometry}
\usepackage{fontspec}
\setmainfont{TeX Gyre Pagella}
\usepackage{enumitem}
\usepackage{hyperref}
\usepackage{xcolor}
\usepackage{fancyvrb}
\usepackage{longtable,booktabs,array}
\usepackage{titlesec}
\titleformat{\section}{\large\bfseries}{}{0em}{}
\titleformat{\subsection}{\normalsize\bfseries}{}{0em}{}
\setlength{\parskip}{0.4em}
\setlength{\parindent}{1.5em}

\title{\textbf{Visualization Architecture}\\
  \large\texttt{ledger-of-meluhha.html} --- Single-File Dashboard}
\author{Third Buyer Advisory}
\date{April 2026}

\begin{document}
\maketitle

%%% ===================================================================
\section{Delivery model}

A single self-contained HTML file. User drops
\texttt{indus\_corpus.db} onto the page. Everything renders
live --- geographic trade map, decoded seals, sign frequency,
weight codebook, radiometric dates. No server. No build chain.
No framework. Just a browser and a database file.

Build chain: \texttt{src/ledger.ts} $\to$ \texttt{npx tsc}
$\to$ \texttt{dist/ledger.js} $\to$ paste into HTML
\texttt{<script>} block.

External dependencies loaded from CDN at runtime:
sql.js (SQLite WASM), Leaflet (map tiles + polylines),
D3.js (frequency chart), Google Fonts (Cormorant Garamond,
JetBrains Mono). Map tiles require internet (OpenTopoMap).

%%% ===================================================================
\section{Design language}

\subsection{Visual identity}

\begin{itemize}[nosep]
\item Background: \texttt{\#0f0c07} (near-black, warm)
\item Accent: \texttt{\#b5121b} (TBA red)
\item Primary text: \texttt{\#f0e6d0} (warm sand)
\item Data colour: \texttt{\#c4924a} (clay)
\item Display font: Cormorant Garamond --- narrative,
  conclusions, titles. Weight 300 for headings. Italic for
  emphasis.
\item Data font: JetBrains Mono --- sign numbers, field labels,
  weights, coordinates.
\item Font is epistemological status. Cormorant = interpretation.
  JetBrains Mono = fact.
\end{itemize}

\subsection{Four themes}

Clay (dark default), parchment (light), high contrast
(WCAG AAA), seal (blue-grey). All colours via CSS custom
properties (\texttt{--clay}, \texttt{--sand}, \texttt{--tba-red}).

%%% ===================================================================
\section{Layout --- three columns}

\subsection{Left panel (\texttildelow20\%)}

Decoded seals M-52A and M-148A. Each as a field card:
F1~merchant mark, F2~commodity, F3~weight tier, F4~quantity,
F5~route. Sign numbers in JetBrains Mono. Resolved values
in Cormorant.

Below the cards: Alloy verification badge in TBA red ---
``16 assertions UNSAT at scope 6.''

Radiometric dating panel: top-5 calibrated dates from corpus DB.

\subsection{Centre panel (\texttildelow55\%)}

Leaflet map centred on IVC trade zone. OpenTopoMap tiles.
22~archaeological sites as circle markers. 13~trade route
polylines as quadratic Bezier curves. Width proportional to
relative trade volume.

Period annotation top-left: ``c.\ 2600--1900 BCE --- Bronze
Age Trade Network'' in JetBrains Mono, dim.

\subsection{Right panel (\texttildelow25\%)}

Three panels stacked:

\begin{enumerate}[nosep]
\item Sign frequency cumulative curve (D3). Claim: ``67 signs
  carry 80\% of all inscriptions. A codebook, not a phonology.''
\item Weight codebook horizontal bars --- binary series then
  decimal. 11~tiers from base unit 0.856\,g. Claim: ``Zero not
  invented until 628 CE. The weight object is the digit.''
\item Commodity manifest table --- sign, commodity, route.
  Click a row to filter matching routes on the map. Active
  row highlighted.
\end{enumerate}

%%% ===================================================================
\section{Width encoding --- trade routes}

Route stroke-width proportional to relative trade volume:

\begin{center}\small
\begin{tabular}{@{}lrl@{}}
\toprule
\textbf{Route} & \textbf{Width} & \textbf{Basis} \\
\midrule
Jar goods $\to$ Mesopotamia & 8\,px & S-342 = 10\% corpus \\
Textiles $\to$ Mesopotamia & 6\,px & Primary bulk export \\
Copper-bronze $\to$ Magan & 5\,px & Khetri to Oman \\
Carnelian $\to$ Dilmun & 4\,px & Luxury export \\
Iron: TN $\to$ IVC & 3\,px & Fish sign supply corridor \\
Timber and Ivory & 2\,px & Sporadic high-value \\
Internal & 1--2\,px & Indus corridor \\
\bottomrule
\end{tabular}
\end{center}

Routes render before sites (z-order: routes under, sites on top).
Polyline refs stored in \texttt{routeLines[]} for commodity
filtering.

%%% ===================================================================
\section{Codebook embedding}

The codebook tables (\texttt{sign\_role}, \texttt{commodity},
\texttt{route}) are embedded as JavaScript \texttt{Map}
constants in \texttt{src/ledger.ts}. These must match
\texttt{IndusCodebookTypes.fsx} exactly --- same sign IDs,
same codes, same values. Sign IDs are Mahadevan numbers.

Weight application labels must match
\texttt{indus\_seed\_codebook.fsx} verbatim. No abbreviations.

\begin{center}\small
\begin{tabular}{@{}lr@{}}
\toprule
\textbf{Embedded map} & \textbf{Entries} \\
\midrule
\texttt{signRoleMap} & 28 \\
\texttt{commodityMap} & 8 \\
\texttt{routeMap} & 5 \\
Weights array & 11 tiers \\
\bottomrule
\end{tabular}
\end{center}

%%% ===================================================================
\section{Seal decode in the browser}

The dashboard decodes seals by parsing Parpola sign IDs from
\texttt{cisi\_inscription} rows in the dropped corpus DB,
then looking up the Mahadevan number in the embedded
\texttt{signRoleMap}. The seal IDs must match the paper
(currently M-52A and M-148A).

Resolution chain:
\begin{enumerate}[nosep]
\item \texttt{cisi\_inscription} $\to$ Parpola sign list
\item Parpola number $\to$ \texttt{signRoleMap} (Mahadevan keyed)
\item \texttt{role=commodity} $\to$ \texttt{commodityMap}
\item \texttt{role=terminal} $\to$ \texttt{routeMap}
\end{enumerate}

%%% ===================================================================
\section{Interactions}

\begin{itemize}[nosep]
\item Hover route: popup with commodity name, note, width.
\item Hover site: popup with name, period, coordinates.
\item Click commodity table row: highlight matching routes
  (non-matching dim to 15\% opacity). Click again to deselect.
\item Route mouseover: weight +2, opacity 1.0. Mouseout
  restores (respects commodity filter state).
\end{itemize}

%%% ===================================================================
\section{Two states}

\subsection{State 1 --- Landing}

Full-screen dark field. TBA wordmark top-left linked to
\texttt{thirdbuyeradvisory.org}. Title centred: ``The Ledger
of Meluhha'' in Cormorant. Drop zone with dashed TBA-red
border. Drop or click to load \texttt{.db} file.

\subsection{State 2 --- Dashboard}

Crossfade from landing (600\,ms opacity transition).
Dashboard \texttt{display:block}. Leaflet \texttt{invalidateSize}
at 100\,ms. All panels render from live DB queries.

%%% ===================================================================
\section{Data flow}

Database file dropped $\to$ \texttt{FileReader} $\to$
\texttt{ArrayBuffer} $\to$ sql.js WASM opens in memory $\to$
queries execute against live schema $\to$ Leaflet + D3 render.

No intermediate files. No server. No JSON export.
The HTML file is the artifact.

%%% ===================================================================
\section{Zenodo artifact}

\begin{Verbatim}[fontsize=\small,frame=single]
ledger_of_meluhha.pdf        — paper (22 pages)
ledger_of_meluhha_overleaf.zip — tex + citations.lua
ledger_of_meluhha_dashboard.zip:
  ledger-of-meluhha.html     — single-file dashboard
  indus_corpus.db            — Alloy-verified corpus
\end{Verbatim}

%%% ===================================================================
\section{Data acquisition roadmap}

Each dataset unlocks more decoded seals on the map:

\begin{enumerate}[nosep]
\item ICIT database (Wells and Fuls, TU Berlin) --- unlocks
  18 pending TN-IVC parallels.
\item Ratnagar (2003) Mesopotamian findspot table --- unlocks
  the falsification test (sign distribution at export sites).
\item Lothal context data from Rao (1973) --- unlocks BC-1
  confirmation (warehouse vs residential context).
\item CISI volumes 2--3 digitisation --- long term, full
  Harappa/Kalibangan corpus.
\end{enumerate}

New corpus data = one command:
\texttt{cmake --build build --target corpus}.
No script, codebook, paper, or dashboard changes needed.

\end{document}
